### 2020-08-30 paper.Rmd
Manuscript edits

---
title: Executive Constraints and Civil Conflict Onset
author:
    name: J.L.A. Krusell
    affiliation: Department of Peace and Conflict Research
    supervisor: Kristine Eck
    degree: "Master's Thesis"
date: "Spring 2020"
titlelogo: |
    ```{=latex}
    assets/UU_logo_CMYK-eps-converted-to.pdf
    ```
wordcount: 15000
suppress-bibliography: false
bibliography: library.bib
csl: assets/chicago-author-date.csl
link-citations: true
documentclass: scrreprt
classoption: paper=letter
geometry:
    - top=30mm
    - heightrounded
fontsize: 12pt
linestretch: 1.5
header-includes:
    - \usepackage{amsmath}
    - \usepackage{etoolbox}
    - \usepackage{lualatex-math}
    - \usepackage{unicode-math}
    - \appto\Shaded{\singlespacing}
    - \BeforeBeginEnvironment{align*}{\begin{singlespace}\vspace*{-\baselineskip}}
    - \AfterEndEnvironment{align*}{\end{singlespace}\noindent\ignorespaces}
    - \counterwithout{table}{chapter}
    - \counterwithout{figure}{chapter}
    - \usepackage[font={small}]{caption}
output:
    pdf_document:
        fig_caption: true
        keep_tex: true
        latex_engine: lualatex
        toc: true
        toc_depth: 2
        pandoc_args: ["--template", "assets/thesis_template.latex",
                      "--syntax-definition", "assets/stan.xml",
                      "--highlight-style", "kate",
                      "--lua-filter", "utils/text_refs/text_refs.lua"]
    html_document:
        smart: true
        toc: true
abstract: |
    **Abstract**: Do institutional constraints on a regime's executive decrease the
    likelihood of civil conflict onset? An unconstrained executive is free
    to loot state resources, undermining the state's capacity to
    effectively deal with nascent rebellions, and may be more likely to
    engage in violent repression, especially in the form of indiscriminate
    violence. This can encourage political grievances, lead to a loss of
    legitimacy for the regime, and provide the opportunity for rebel
    groups to attract new members by offering protection against
    government violence. Furthermore, the lack of guaranteed checks to
    executive power may incentivise actors to avoid bargained solutions if
    they fear future extralegal retributions. Taken together, where there
    is a lack of effective constraints against the executive there may be
    an increased risk for the onset of civil conflict. To test this
    proposition I employ a bayesian latent variable model which estimates
    executive constraint as a latent factor derived from several manifest
    variables and relates it to a binary measure of civil conflict onset
    through a logistic regression.  The primary finding is that there is a
    negative relationship such that the predicted probability for civil
    conflict is lower where there exists higher levels of executive
    constraints. However, this is conditional on the level of GDP per
    capita; for low-income country-years the relationship between
    executive constraints and civil conflict onset is indeterminate
    possibly because it is easier to recruit and mobilize fighters
    regardless of the overall level of executive constraints. The issue of
    poor model fit means that the presented results should be considered
    tentative at best.
---

[//]: # Setup R environment for entire doc
```{r include = F, results = "hide"}
library(corrplot)
library(dplyr)
library(extraDistr)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(magrittr)
library(kableExtra)
library(knitr)
library(parallel)
library(precrec)
library(sf)
library(tidyr)

# Load last to ensure nothing overwrites our namespace
library(thesis.utils)

options(knitr.kable.NA = ".")
opts_chunk$set(echo = F, message = F, warning = F,
               fig.width = 5, fig.height = 3, cache = T)
```

# Introduction {-}

```{r intro-descriptive-stats}
# Create initial descriptive stats for introduction
ucdp <- readRDS("./data/raw/UcdpPrioConflict_v19_1.rds")

total_conflicts <- n_distinct(ucdp$conflict_id)
intrastate_conflicts <- filter(ucdp, type_of_conflict %in% 3:4) %$%
    n_distinct(conflict_id)
intrastate_percent <- 100 * intrastate_conflicts / total_conflicts

n_states <- filter(ucdp, type_of_conflict %in% 3:4) %$%
    n_distinct(gwno_a)

battle_deaths <- readRDS("./data/raw/BattleDeaths_v19_1.rds")
nonstate_deaths <- readRDS("./data/raw/NonState_v19_1.rds")

total_deaths <- sum(battle_deaths$bd_best) + sum(nonstate_deaths$best_fatality_estimate)
civil_deaths <- filter(battle_deaths, type_of_conflict %in% 3:4) %$%
    sum(bd_best)

civil_deaths_percent <- 100 * civil_deaths / total_deaths
```

Since the end of World War II, civil conflict has been the most
prevalent form of state-based conflict. Of the
`r fmt(total_conflicts)` conflicts recorded by the UCDP/PRIO Armed
Conflict Database v19.1 [@Gleditsch2002; @Pettersson2019] for the
years covering 1946 until 2018, `r fmt(intrastate_conflicts)` of
those, or approximately `r fmt(intrastate_percent, 1)` percent, are
classified as intrastate. These civil conflicts are distributed across
`r fmt(n_states)` countries, more than half of which having
experienced at least one major civil war with 1,000 battle deaths at a
minimum.  Where data is available, civil conflicts are also
empirically among the deadliest. Out of `r fmt(total_deaths)`
cumulative battles deaths from organized violence recorded since 1989
[@Pettersson2019; @Eck2007; @Sundberg2012], `r
fmt(civil_deaths_percent, 1)` percent are the result of civil
conflict.

Given the impact of civil conflict and the increasing availability of
conflict related data, the last several decades have witnessed a
veritable explosion in the number of quantitative studies
investigating the causes of civil conflict [@Blattman2017]. Scholars
have proposed numerous theories ranging from motivation-based
arguments --- such as, for example, the role of horizontal
inequalities and ethnic identities [@Cederman2011] --- to
opportunity-based explanations centered on economic factors
[@Collier2004; @Collier2009] as well as the political and structural
conditions underlying state weakness [@Hegre2001; @Fearon2003]. This
latter strain of research in particular has spurred extensive work
into examining the distinguishing features of weak states and whether
political institutions play a significant role in the risk of civil
conflict. Several studies have argued in favor of institutions and
have investigated, for example, aspects of governance, institutional
quality, and opportunities for political participation as factors
influencing the risk of civil conflict [@Fjelde2009; @Taydas2010;
@Hegre2015a; @Bartusevic2018].

This paper extends that work by taking a disaggregated approach to
studying political institutions by focusing specifically on the effect
of executive constraints on the risk of civil conflict onset. The
motivation to focus on executive constraints is two-fold: first, the
executive is often times one of the key decision-makers in a state's
regime. Institutions that provide a check against the types of actions
which may exacerbate the risk of rebellion should be relevant to the
study of civil conflict. Such an approach can further our
understanding of how specific political institutions may shape the
incentives for violence by affecting state behaviour. Second, the link
between executive constraints and civil conflict is relatively
understudied. The previous literature that exists is fairly limited
and contradictory in terms of results.

The principle finding of this paper is that there is tentative
evidence that executive constraints do matter and that there is a
negative relationship between the latent level of constraints on the
executive and the risk of civil conflict onset. However, this
relationship largely disappears for low-income country-years when an
interaction term between executive constraints and gross domestic
product (GDP) per capita is included. If the primary theoretical
mechanism involves the impact of executive constraints on rebel
mobilization --- an unchecked executive may be more likely to engage
in indiscriminate violence which could lead to a loss of legitimacy as
well as the ability of rebel groups to offer local security guarantees
to attract recruits --- then the incentives for violence in low-income
country-years due to say economic factors may be so great that it does
not matter how constrained the executive may or may not be.

We should not be overconfident with these results, however. Similar to
most other studies examining conflict onset, the presented models are
a poor fit to the underlying data. Modeling civil conflict is a
difficult task, especially when trying to capture the dynamics of
conflict with a standard time series cross-country regression. The
results will require verification by additional studies using
alternative methodological approaches.

In what follows, a deeper justification for why political institutions
are relevant to the study of civil conflict together with an overview
of related empirical literature will be presented. The theoretical
mechanisms linking executive constraints to conflict onset will then
be discussed followed by an empirical analysis using a bayesian latent
variable model to estimate executive constraints from the basis of
several observed manifest variables. This latent measure will then be
used to predict civil conflict onset using a cross-country time series
logistic regression.

# Theoretical Background {-}

## Political Institutions and Civil Conflict

This paper takes as a starting premise a definition of institutions as
being "systems of established and embedded social rules that structure
social interactions" [@Hodgson2006]. More informally, we can consider
them to be the "rules of the game", encompassing, for example, codified
laws, norms, and shared understandings, all of which can constrain or
enable behaviour by influencing the perceptions and preferences of
actors through incentives and deterrents.[^institutions] Furthermore,
over time, the habits formed from these systems create expectations of
stable patterns of behaviour thereby further shaping the interactions
between actors by reducing uncertainty [@North1990]. Although there is
a wide variety of institutions, the present focus will be on the type
of institutions which govern political behaviour.

[^institutions]: None of which is meant to imply that institutions
    stand external to actor perceptions and preferences as unchanging
    features. Rather, institutions are social constructs subject to
    changing normative values and able to be reinforced or undermined
    by behavioural feedback loops. This latter point is especially
    interesting within the context of civil conflict considering how
    the development of various institutions and conflict incidence can
    by jointly determined; however, it remains outside the scope of
    this paper.

This exposition is relevant to the study of civil conflict for both
motivation-based and opportunity-based explanatory accounts. To the
extent that the onset of civil conflict is a longer process featuring
waves of escalation and de-escalation rather than a single one-off
event [@Florea2012; @Sambanis2004b], institutions are a key influence
on the constitutive interactions between state and non-state
actors. For example, they can shift the opportunity costs of violence
altering the mobilization capacity of would be rebel
groups. Institutions can also determine the extent to which the state
is seen as accountable to and representative of the larger population,
thereby influencing the legitimacy of anti-government violence. In
addition, institutional quality can shape the capacity of the state to
suppress or co-opt potential opposition. Although the preceding
examples are far from forming an exhaustive list, they illustrate the
number of different ways in which institutions can enter into
theoretical discussions about civil conflict.

Unsurprisingly, insofar as the role of institutions in shaping the
likelihood of civil conflict has been examined by previous
literature, the focus has largely been on the broader notion of
democracy, which at a minimum is a system regulating access to
political power through popular representation determined by free and
competitive elections [@Schumpeter1962]. Democracy serves as a proxy
for a number of normatively desirable institutions and has been
hypothesised to have a pacifying domestic effect according to two
predominant mechanisms. The first is centered on the concept of
relative deprivation, or inequalities in value expectations and value
capabilities, as a primary source of discontent that can motivate
collective violence [@Gurr1970]. [^gurr] Democracy is thought to
decrease the amount of political discontent by providing avenues for
free expression and adjudicating group conflicts through non-violent
institutional means --- for example, through the regular occurrence of
free and fair elections [@Hegre2014]. The second mechanism assumes
that violent conflict is costly and that there exists a bargained
solution preferable to fighting, assuming that actors are rational and
able to internalize the cost of violence [@Fearon1995]. One of the
reasons why a bargained solution may be impossible is the risk of
future shifts in the distribution of power. More specifically, a
future change in favor of the relative strength of one of the parties
can provide an incentive to later renege. This influences the present
degree to which concessions can be offered and parties are willing to
credibly commit [@Powell2006]. Within the domestic sphere, democracy
offers a possible solution by formalising the regular contest of power
with full enfranchisement, thereby providing a *de jure* guarantee for
the balance of power [@Origins2006], and should, likewise, decrease
the likelihood of civil conflict.

[^gurr]: Gurr explains value expectations as being that which we think
    we are entitled to and value capabilities as that which we think
    we are capable of attaining. As Gurr then states, relative
    deprivation is the perceived discrepancy between the two. The link
    to collective violence, however, is conditional on the discontent
    being politicized, by which it is meant that the source of the
    discrepancy is identified and a normative and utilitarian
    justification is given for violent action, before finally being
    actualized.  Quantitative studies tend to focus on the first
    stage, the propensity for political regimes to generate
    discontent, given the difficultly of measuring the factors that
    may affect politicisation and actualisation.

Despite the theoretical basis linking democracy to reduced risks of
civil conflict, the empirical results tend to be mixed. @Hegre2001,
for example, find that democracies are at a lower risk for civil
conflict; however, no more so than fully authoritarian
states. Instead, anocratic regimes in between democracy and
authoritarianism are considered to be the most prone to civil conflict
onset. The explanatory logic is that both democratic and fully
authoritarian states tend to be consolidated regimes. In the case of
the former, this means the presence of institutions conducive to
conflict resolution such that the opportunity for rebellion may be
present, given the availability of civil liberties conducive to
organizing, but there is a notable lack in motivation for collective
violence. For the latter, authoritarian regimes may produce a
significant amount of political discontent; however, the opportunity
is lacking due to the state's capacity and willingness to violently
repress dissent. Anocratic regimes are by their very nature
transitory, partly open and partly repressive, encouraging political
discontent against the state without the capacity to either
accommodate or suppress opposition.

In contrast, studies emphasizing the importance of political and
economic opportunities for rebellion in understanding when and where
conflicts break out, such as @Fearon2003 and @Collier2009, find
democracy to be insignificant. According to their theoretical
expositions, democracy should be associated with higher levels of
political rights, respect for ethnic and religious minorities, and
less repression, thus being a proxy for the latent level of political
discontent against the state. However, the argument follows that even
if the number of political grievances vary across regime types, they
are still ubiquitous enough to be a poor explanation for why conflict
breaks out in only a relatively small subset of countries at any given
point in time. Instead, the determining factors for conflict onset are
the structural conditions conducive to rebellion.[^opportunities]

[^opportunities]: For Fearon and Laitin these structural conditions
    are related to the features which make rebellion militarily
    viable.  Specifically, there's a higher risk of conflict onset
    where state capacity is weak, principally understood in terms of
    state revenue and its impact on policing and counterinsurgency
    operations; large impoverished populations; and insurgents having
    a terrain advantage. Collier *et al* start from the premise that
    war is expensive and the primary impediment to a rebellion is its
    financial feasibility relative to the strength of the state. On
    the basis of their results they go so far as provide support to
    the proposition, "where a rebellion is financially and militarily
    feasible it will occur" [@Collier2009]. Both approaches, however,
    reach the same conclusion concerning the empirical evidence
    surrounding the role of democracy.

To provide a more nuanced picture, an ongoing strain of research has
been dedicated to disaggregating measures of democracy into either
finer regime type categories or specific measures of individual
political institutions. The theories offered tend to follow a general
formula whereby formal lack or poor quality of some institution(s),
measured either directly or proxied by regime type, undermines the
capacity of the state to prevent armed insurgency and/or incentivises
opposition violence by restricting the opportunities for political
participation and increasing political discontent. For example,
@Carey2007 examines the executive selection process of Sub-Saharan
Africa and argues that executive elections, even when they are of poor
quality, reduce the risk of large-scale violent dissent. @Fjelde2010
breaks down authoritarian regime types in Africa attributing the
significant variation in conflict risk to the differing institutional
configurations affecting a regime's ability to coerce or co-opt the
opposition. The conclusion is that previous indeterminate results
regarding the link between political institutions and conflict risk
may have been hamstrung by aggregate measures which poorly reflect the
within-group variation of political regimes.

In a larger sample including democratic polities, @Goldstone2010
create a regime classification scheme based on measures for openness
of executive recruitment and competitiveness of political
participation, as well as factionalism in order to distinguish between
partial democracies. They offer a theory of civil conflict occurring
in countries where elite divisions undermine the capacity of the state
to defeat an insurgency and claim that a relatively parsimonious model
containing their regime classification offers better predictive
accuracy than previous analyses and provides evidence that political
institutions are a dominant driving factor for both political
instability and civil conflict. Finally, @Bartusevic2018 construct a
regime typology focused specifically on electoral features, which are
argued form the essential core of the concept of democracy. Where
there are constraints on electoral contestation, the authors contend
that such constraints restrict the opportunities for non-violent
political participation thereby legitimising anti-government violence
and self-selecting recruits into rebel groups willing to engage in
armed conflict. Evidence is presented which the authors claim supports
not just the significance of regime type, but a negative monotonic
relationship between democracy and the risk of civil conflict.

Studies utilising individual measures beyond regime classification
have been largely focused on the concept of governance. For instance,
@Taydas2010 examine institutional quality as a means to capturing a
regime's governing capacity which is theorised to be linked to
conflict onset both in terms of political discontent --- a regime's
inability to govern effectively and respond to public needs leads to
grievances and a loss in legitimacy --- and political opportunities
for rebellion --- poor institutional quality undermines the state's
ability to handle a nascent, violent uprising. Specifically, they
assess corruption, rule of law, and bureaucratic quality and present
evidence that all three are negatively related to civil conflict
onset. Similarly, @Hegre2015a argue that the presence of formal
representative institutions does not fully capture the extent to which
a regime is capable of realizing policy goals benefiting the
public. Poor quality of government can therefore undermine the
legitimacy of a regime when citizens identify the state as responsible
for relative deprivation. In their empirical analysis, the authors
separate informal institutions related to good governance from formal
democratic institutions and find both important for reducing the
incidence of recurring civil conflict.

Taken together, this recent literature on civil conflict onset points
to the ongoing importance of political institutions. There remains,
however, significant work to be done identifying the causal mechanisms
and effects of specific institutions, especially in terms of further
disaggregating the concept of democracy beyond ever finer
classifications of regime type [@ari2016]. Furthermore, to the extent
that previous research has mostly focused on the effects of
institutions on would-be rebel motivation and/or political opportunity
structures, there is also a gap in terms of empirical analyses on how
institutions also shape the behaviour of the state in the lead up to a
civil conflict. Even in strictly economic explanations of civil
conflict where explanatory primacy is given to either greed-based
motivation or economic conditions conducive to rebellion, the state
is a key actor. How a state responds to dissent structures the
opportunity costs of violence as well as the salience of calls to
mobilization by rebel leaders. This latter point is especially
relevant in linking regime legitimacy to conflict since public
perception is shaped not simply by a regime's formal structure, but by
its conduct in repeated interactions with the public. To this end, it
is imperative to examine the types of institutions which exacerbate
the risk of conflict onset by enabling or constraining certain actions
by the state.

There are of course a large variety of political institutions for
which the presence and quality can influence the behaviour of the
state.  Given the importance of the executive as a central authority,
however, especially in times of political crisis, institutions which
affect the executive's decision making should be pertinent. Chief
among these is the idea of executive constraints --- more formally,
the institutional constraints by other government bodies on the
executive's ability to abuse power and act unilaterally. Executive
constraints can be considered a subset of horizontal accountability
and imply the existence of state agents --- such as the legislature or
the judiciary --- with both the legal mandate and willingness to
oversee or sanction the executive thus serving as deterrents to the
forms of authoritarian excesses in executive power which may be linked
to civil conflict. [^ODonnell]

[^ODonnell]: This is adapted from the definition of horizontal
    accountability offered by @ODonnell1998 who further argues that
    horizontal accountability can potentially be a distinguishing
    feature between consolidated democracies and newly democratic
    states, the latter of which may have free and open elections, but
    are considerably weaker in upholding liberal notions of individual
    freedom against state power and republican notions of
    representation. In this way, executive constraints can possibly
    aid in discriminating between the types of partially democratized
    regimes which may be at a heightened risk for civil conflict.

The final motivation for examining executive constraints is that in
terms of previous research, the empirical evidence is
conflicted. @Getmansky2013, for example, offers a theory of executive
constraints playing a role in risk aversion in democratic polities.
When the executive is constrained by other government institutions a
regime may be unwilling to resort to force against a growing
insurgency and may instead seek a peaceful solution to avoid
fighting.[^weakness] They use a dichotomized measure of XCONST,
"Constraint on Chief Executive", from Polity IV [@Marshall2019] in
order to test this theory, finding their measure to be insignificant
in predicting the outbreak of insurgencies.

[^weakness]: If the theory were to hold, it is unclear if risk aversion
    really translates to a lower incidence rate of civil conflict
    since such regimes may be subject to opportunistic rebel groups
    ready to exploit precisely those weaknesses. In other words,
    democratic regimes may actively attempt to avoid armed conflict;
    however, they can still be forced into such situations by rebel
    groups.

In another study, @Fjelde2009 put forth a theory tying executive
constraints to public provisioning. Increased government expenditures
in the presence of strong institutional checks against executive
corruption is thought to decrease the risk of civil conflict by
reducing potential grievances against the state. In contrast to
Getmansky, the authors find a significant negative effect for an
interaction term between executive constraints and government
expenditures when also using XCONST from Polity IV, although the
regression slope for the executive constraints term is insignificant.

In both studies, using an ordinal measure, and in the case of
@Getmansky2013 a binary measure, for executive constraints risks
losing a significant amount of information concerning the varying
degrees of effectiveness in the constraints against executives among
different countries. A possible improvement would be the use of a
continuous measure to be more able to capture the quality of
constraints and better adjudicate whether there exists a measurable
effect on civil conflict onset.

## Executive Constraints and Civil Conflict

There are three general pathways which can theoretically link the
level of executive constraints to the onset of civil conflict:
executive looting and state weakness, credible commitments to
bargained solutions, and the impact of violent repression by the state
on rebel mobilization. In the case of the first, the absence of
effective constraints enables leaders to loot state resources
[@Blattman2017], diverting funds for the purposes of personal
rent-seeking and corruption, which can undermine the administrative
and military capacity of the state to combat a nascent
insurgency. Furthermore, such environments allow leaders to pay
selective benefits to close supporters at the expense of wider
redistributive policies. To the extent that the provisioning of public
goods may ameliorate grievances against the state and co-opt potential
opposition, a reduced risk for civil conflict could be expected where
there is a high level of redistributive public spending and the
presence of quality institutions able to check the ability of the
executive to engage in cronyism [@Fjelde2009]. Finally, the lack of
accountability stemming from missing executive constraints could make
the state an attractive prize to predatory rebel groups; however,
prior empirical evidence lends little credence to this mechanism
[@Bazzi2014].

The second pathway relates to the willingness of both the regime and
opposition to credibly commit to bargained solutions to avoid
conflict. For the opposition, the presence of government institutions
able to hold the executive to account with the threat, for example, of
investigations, sanctions, or court rulings reduces the uncertainty
over whether the current leader will later renege. From the
perspective of the regime, executive constraints serve to limit future
violent retributions if oppositional groups were to take power. This
latter point is especially salient in situations where a leader has
previously relied on violent repression to stay in power and fears
extralegal punishment. It should be noted, however, that in such
authoritarian settings the antecedent lack of effective executive
constraints does not explain why the leader decides to negotiate in
the first place. But, insofar as the bargained solution includes a
transition of power with democratic reforms, and possibly even
immunity for the current members of the regime, the inclusion of
future checks against the executive serves as a guarantee for the
current agreement thereby increasing the incentive to commit.

The final pathway involves the impact of executive constraints on
shaping the ability of rebel groups to recruit new members. In order
to mount an effective challenge to the state, rebel leaders need to be
able to mobilize fighters while avoiding defections. However,
participating in an armed rebellion is a costly endeavour. Non-elites
face both the immediate risk of fighting as well as the possibility of
future reprisals by the government. Furthermore, given that most civil
conflicts are protracted events against an often times much stronger
opponent, the short-term utility from participation is relatively
low. Thus, rebel leaders must be able to offer proportionate selective
benefits to offset the associated costs of fighting
[@Mason1996]. These benefits can include not only economic payoffs,
but also protection from government violence. When the government
engages in violent repression, especially in the form of
indiscriminate violence, the cost of non-participation is
raised. Rebel groups can provide local security guarantees against
government reprisals such that the opportunity cost of fighting is
lowered, thereby incentivizing participation [@Regan2005]. However,
this is predicated on the government lacking the capacity to fully
eradicate any rebel group. Where that capacity exists the state is
free to engage in reprisals.

It is at this point where the concept of executive constraints then
enters into the calculus. Leaders in nondemocratic states are more
likely to engage in violent repression as a response to internal
dissent [@Davenport2007a]. Executive constraints provide an
accountability mechanism deterring state violence by the leader
against their own citizens. Thus, where there are strong and effective
constraints on the executive, the opportunity cost for joining the
rebel side should be higher since the lack of indiscriminate violence
makes non-participation an attractive alternative. Rebel leaders will
then have to rely more heavily on economic selective benefits
[@Regan2005], creating a higher barrier for mobilization and
decreasing the risk of civil conflict onset. It should be noted
though, that where this relationship may not hold is in situations of
low intensity insurgencies involving predatory rebel groups solely
motivated by rent-seeking, or resource extraction, rather than
secession or government overthrow. The threshold for mobilization in
such scenarios is significantly lower than in full scale civil wars
since such rebel groups are not actively challenging the state for
control, meaning that rebel groups can initiate a conflict regardless
of the current level of state repression.

Taken together, none of the aforementioned mechanisms are mutually
exclusive, and it is beyond the scope of the present paper to properly
adjudicate between them; however, they do converge on a single
testable proposition. Namely that higher levels of executive
constraints should be associated with a lowered risk of civil conflict
onset. This proposition can be further extended: if state weakness is
a primary driver of civil conflict and the relative capacity of the
state is effected by the levels of executive constraints, as depicted
in the first theoretical pathway regarding executive looting, then
state capacity could have a mediating affect on the relationship
between executive constraints and civil conflict onset. Another
extension would be if executive constraints affect the mobilization
potential for rebel groups, then there could also be a confounding
effect from the relative threshold for mobilization. Finally, the
relationship could still be confounded by the overall opportunity for
rebellion --- the degree of executive constraints may not be
particularly relevant in relatively stable countries lacking the
conditions conducive to carrying out a rebellion. Regardless of
conditionalities, contrary results would indicate either that
executive constraints are indeterminate to the risk of civil conflict
or possibly that there is even a positive relationship to conflict
onset since constrained executives may be less able to effectively
deal with violent challenges due to risk aversion to violence and/or
from the slowdown of having to seek approval from a plurality of
stakeholders before committing to decisive action.

One final caveat, hitherto the relationship of executive constraints
to the notion of vertical accountability --- constraints on leaders by
free and open multiparty elections as well as an independent and
critical media --- has been left unspecified. This is not to diminish
the role of elections or the media; indeed, vertical accountability
can plausibly play a role in all of the aforementioned mechanisms. For
example, studies on state repression have emphasised the importance of
non-elites in being able to vote elected offices out of office as an
effective deterrent against negative sanctioning by leaders
[@BuenoDeMesquita2005; @Davenport2007]. But, elections occur only
intermittently and it is not an automatic given that voters will
punish an incumbent, especially when that repression is against an
oppressed minority community. Furthermore, while the media can expose
wrongdoing by government officials, legal punishment can only flow
through government bodies [@ODonnell1998]. Thus, executive constraints
are interesting in their own right and vertical accountability should
not confound any potential negative relationship to civil conflict
onset.

# Empirics {-}

## Research Design

To examine the relationship between civil conflict onset and executive
constraints, I construct a dataset of relevant variables and run a
bayesian latent variable model with a measurement portion
estimating executive constraints as a latent construct and a
multilevel logistic regression relating the latent construct to the
observed outcome, civil conflict.

The use of bayesian methods is particularly advantageous in this
setting since the constructed dataset is comprised of the entire
finite population of countries within a non-random subset of years
given only non-stochastic missingness. Traditional frequentist
estimators are inappropriate due to the assumption that the observed
data is drawn from a random, repeatable generative mechanism. This is
readily apparent when attempting to interpret a frequentist confidence
interval in the present context. A standard confidence interval for a
selected level of significance, $\alpha$, reports only that $(1 -
\alpha) 100 \%$ of the calculated confidence intervals from additional
inferences done on repeated samples are expected to contain the true
population value [@Western1994]. However, for the current analysis the
desired target is really to be able to calculate directly the
probability that the current uncertainty interval contains the true
value given the data and the model specification. Bayesianism offers a
solution by casting probability as degrees of belief in some event
wherein inference is then the process of updating those prior beliefs
in light of new data. When using bayesian methods, parameters such as
regression coefficients are considered to be random, meaning that
prior beliefs in the form of probability distributions can be assigned
and posterior probability distributions estimated from observed data
[@Western1999].

In what follows, I will first describe the regression portion of the
model before delving into the independent and dependent variables,
including the operationalizations and data sources as well as the
measurement model for executive constraints.[^model]

[^model]: The model code is written in Stan [@Carpenter2017] and was
    run using the `CmdStan` interface version 2.23.0 [@Stan2020] with
    four parallel chains at 2000 iterations each with a warmup
    of 1000. Convergence was assessed according to the potential scale
    reduction factor $\widehat{R}$ and effective sample size (ESS)
    [@Vehtari2019] in addition to standard Stan diagnostics. All
    pre/post-processing was done using R 3.6.2 [@Core2019]. The Stan
    model code is available in [Appendix A](#appendix-a) and the full
    project in the form of a reproducible pipeline, including the
    Rmarkdown source for this manuscript, is available at
    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).
    Finally, a simulation analysis assessing the model and it's
    ability to recover simulated true values is available in [Appendix
    B](#appendix-b).

[//]: # Load merged dataset for descriptive stats in following sections

```{r results = "hide"}
merged.df <- readRDS("data/merged_data.rds")
```

### Regression Model

I model civil conflict onset as a multilevel regression where onset is
a function of a set of predictors, including executive constraints and
an array of control variables, together with several varying
intercepts. The units of analysis are all independent country-years
from 1952 until 2013, as determined by data availability, with *I*
countries indexed by $i = 1 \dotsc I$ and *T* years indexed by $t = 1
\dotsc T$. The dependent variable, conflict onset, is a Bernoulli
distributed binary variable where 1 denotes the onset of at least one
new civil conflict in a given country-year. The chosen link function
transforming the linear regression output into the probability of
conflict onset, $\nu_{it}$, is the inverse logit such that $\nu_{it}
\in [0, 1]$. [^invlogit] The choice of a logit link function avoids
the truncation necessary for a linear probability model. Being a
sigmoid function it also ensures that the marginal contributions from
the predictors are reduced at the tails for higher and lower
probability values. That is to say the added value of a one unit
increase in a predictor is higher when the predicted probability is
already at 0.5 at the boundary between a positive and negative outcome
versus when it is at 0.95.

[^invlogit]: Technically, the probabilities in a logit function are on
    the open interval, $(0, 1)$, but due to overflow/underflow in
    floating point calculations done by computers the endpoints can be
    observed.

The full regression specification is then as follows:
\begin{align*}
  y_{it} & \sim \mathsf{Bernoulli}(\nu_{it}) \\
  \nu_{it} & = \mathsf{logit^{-1}}(\alpha + X_{it-1} \beta +
                 \sum_{k=1}^K \zeta^k_{j_{it}})
\end{align*}

where **X** is the predictors including the latent measure for
executive constraints, $\beta$ the vector of regression slopes, and
$\Zeta$ the *K* sets of varying intercepts indexed by $j_k$. All
predictors are lagged by one year to control for time endogeneity and
standardized by subtracting the mean and dividing by the standard
deviation to aid in computation as well as comparison of the posterior
$\beta$ estimates.

For the present analysis there are three sets of varying intercepts,
or random effects, included in the model. This includes two partially
pooled grouping factors, year and country, to account for
cross-sectional unmodeled heterogeneity between countries and over
time [@Shor2007]. Also included is the number of peace years since the
last civil conflict incidence per country to control for the lingering
endogenous effects of violence on the likelihood that additional
conflict breaks out.

I place weakly informative priors on the population intercept and
slope parameters to allow the posterior estimates to be largely
determined by the observed data while avoiding the more pathological
values and hard constraints that can arise from uninformative
priors. Given that the input data has been standardized, this
translates to:
\begin{align*}
    \alpha & \sim \mathsf{Normal}(0, 5) \\
    \beta & \sim \mathsf{Normal}(0, 2.5)
\end{align*}

For the varying intercepts modeling year and country specific
variability, I use hierarchical priors to provide a regularizing
effect against overfitting [@Gelman2007].
\begin{align*}
    \zeta^{year}_t & \sim \mathsf{Normal}(0, \sigma^{year}) \\
    \zeta^{country}_i & \sim \mathsf{Normal}(0, \sigma^{country}) \\
    \sigma^{year}, \sigma^{country} & \sim \mathsf{Cauchy_{+}}(0, 1)
\end{align*}

Finally, for peace years I take a non-parametric approach and use
gaussian process priors with an exponentiated quadratic kernel. This
treats the number of peace years as an infinitely continuous dimension
along which the covariance between any two given points is a function
of their euclidean distance. Where that distance is smaller there is a
higher correlation. Compared to other methods for modeling time
dependence this approach avoids the parametric assumption of a
proportional rate in exponential decay functions, the difficulty in
selecting interaction terms to account for non-proportionality for
cubic polynomial approximation, and the somewhat arbitrary selection
of knots for splines [*c.f.* @Raknerud1997; @Carter2010; @Beck1998].

Given the finite set of observed peace years in the dataset, the
specification is as follows:
\begin{align*}
    \zeta^{peace\_years} & \sim \mathsf{MultiNormal}(0, K) \\
    K_{ij} & = \eta^2 \mathsf{exp}(\minus \frac{1}{2 \rho^2} D_{ij}^2)
\end{align*}

where given $P$ unique peace years, $K \in \mathbb{R}^{P \times P}$ is
the covariance matrix, $D \in \mathbb{R}^{P \times P}$ the distance
matrix, and $i$ and $j$ the column and row indexes. The
hyperparameter, $\rho$, is the *length-scale* and corresponds roughly
to the number of years before two observations become
uncorrelated. The hyperparameter, $\eta$, is the *marginal standard
deviation* which determines the overall variation of the output
intercepts [@Rasmussen2005].

I select a weakly informative prior for $\eta$ that is skeptical of
large values given that $\nu$ is on the log-odds scale. For $\rho$ I
use an informative prior that constrains an overwhelming majority of
the prior probability mass to a *length-scale* between 4 and 10, which
provides a strong smoothing effect and aids in computation by avoiding
unreasonably large values far beyond the observed range
[@Betancourt2017].
\begin{align*}
    \eta & \sim \mathsf{Normal_{+}}(0, 1) \\
    \rho & \sim \mathsf{InvGamma}(15.5031, 78.2992)
\end{align*}

### Civil Conflict

The dependent variable, *civil conflict onset*, is taken from the
UCDP/PRIO Armed Conflict Dataset (ACD) v19.1.[^onset_dataset]
According to the dataset, intra-state conflict is defined as a violent
contest over a government or territory between a state and non-state
actor with a minimum threshold of 25 battle deaths within a single
country-year with a coverage from 1946 to 2018.  The data was merged
based on the government side for conflicts labeled as internal or
internationalized internal --- *i.e.* where *type_of_conflict* was
either 3 or 4. Multiple conflicts within the same years were
aggregated to form a binary variable coded as 1 to denote the onset of
at least one new civil conflict and 0 for all other years, including
years of ongoing conflict.[^conflict_type]

[^onset_dataset]: UCDP also makes available a pre-aggregated onset
    dataset for internal and internationalized internal armed
    conflicts. However, there are a number of coded onsets which do
    not correspond to the conflict years listed in the standard
    ACD. For example, China is coded as experiencing episode onsets in
    1969, 1974, 1978, 1983, 1986 with unique conflict starts in 1969
    and 1974 according to the *onset1* and *newconf* variables in the
    onset dataset.  In the ACD, however, China is only coded as being
    engaged in an interstate conflict with Myanmar/Burma (1969),
    Russia (1969), and Vietnam (1974 - 1988, with several gap
    years). Likewise discrepancies exist for a number of other
    countries affecting 80 country-years in total. Since the onset
    dataset does not inlude identifying information on each conflict,
    I use the larger ACD to programmatically calculate the conflict
    onsets.

[^conflict_type]: Extrasystemic conflicts, *i.e.* imperial or colonial
    conflicts, were excluded largely due to lack of data. Notably,
    there is very little easily available economic data for colonial
    territories and it would be undesirable to use the data for the
    colonizing power due to the often wide discrepancies in
    development within colonial empires.

The primary onset dates were determined according to UCDP/PRIO's
episodic definition where a new onset is coded for the same
incompatibility when there is at least a one year pause in the
fighting or drop below the battle death threshold. The primary
rationale is that civil conflicts often go through several distinct
phases of fighting which can make it difficult to accurately code the
end of a conflict. The episodic definition avoids missing onsets from
the re-emergence of previously long dormant conflicts.  The risk
though is the inclusion of added onsets from low intensity conflicts
due simply to intermittent drops in the recorded number of battle
deaths per year. However, the posited theories linking executive
constraints to civil conflict should still hold even in such
scenarios. The incentive structures surrounding political violence are
still relevant to rebel groups even in intermittent years. Groups need
to be able to continue to attract fighters in order to maintain a
credible threat, in which case the level of executive constraints
should still have its theorized effect.

In total there are `r sum(merged.df$episode_onset)` civil conflict onsets in the
assembled dataset according to the episodic definition affecting
`r filter(merged.df, episode_onset == 1) %$% n_distinct(country_name)`
unique countries out of a total of
`r n_distinct(merged.df$country_name)`. Figure \ref{fig:onsets} plots
the proportion of countries experiencing an ongoing civil conflict and
the proportion of countries experiencing new onsets of conflict over
time.

(ref:onsets) Proportions of countries per year experiencing at least
    one ongoing civil conflict and at least one new onset.

```{r fig.cap = "\\label{fig:onsets} (ref:onsets)"}

df <- group_by(merged.df, year) %>%
    summarise(`New Onsets` = sum(episode_onset) / n(),
              `Ongoing Conflict` = sum(ongoing) / n()) %>%
    gather(variable, value, -year)

ggplot(df, aes(year, value, colour = variable)) +
    geom_line() +
    theme_tufte() +
    scale_color_colorblind() +
    theme(legend.title = element_blank()) +
    ylab("Proportion of Countries") +
    xlab("Year")

```

### Measuring Executive Constraints

To measure executive constraints I use data from the Varieties of
Democracy (V-Dem) dataset v9.0 [@Coppedge2019]. V-Dem is a
cross-national time series dataset providing multidimensional,
disaggregated data on democracy and political institutions for over
200 countries with an overall year coverage from 1789 until 2018. The
principal data is drawn from expert coded surveys involving a global
network of over 3,000 coders, most of whom are academics or
professionals working in policy related fields. Individual ratings are
aggregated into latent variable estimates using hierarchical ordinal
item response theory (IRT) models which take into account rater
reliability and varying rater thresholds
[@Pemstein2019]. [^thresholds] The result is over 220 lower level
indicators that are also combined through various aggregation
techniques to produce over 80 indices. Compared to datasets such as
Polity IV, which has traditionally been used in most analyses of civil
conflict, V-Dem provides much more fine grained, nuanced data, and
most importantly, includes measures of uncertainty propagated from the
IRT models. This latter point is crucial since ignoring measurement
error, especially in models with noisy data and small effect sizes as
is often the case in conflict studies, can bias estimates and lead to
overconfident results [@Treier2008; @Marquardt2019].

[^thresholds]: Thresholds relate to how the continuous latent trait
    for a given variable is mapped on to the ordinal answer categories
    presented in the survey. Due to differential item functioning,
    experts may perceive the magnitude of difference between answer
    items differently. For example, for the variable *v2elvotbuy*,
    election vote buying, experts may vary in how they interpret the
    difference between a 2 ("Restricted Vote Buying") and a 3 ("Almost
    No Vote Buying"). These two concepts, rater reliability and
    varying rater thresholds, correspond respectively to the
    *discrimination* and *difficulty* parameters in classical IRT
    models.

For the purposes of this paper I estimate an index measuring executive
constraints from a factor analysis of multiple manifest variables
drawn from the V-Dem dataset. The aim is to capture the extent to
which the executive is *de facto* constrained by other government
institutions --- more specifically, the legislature, judiciary, and
constitution --- rather than the *de jure* powers of any one
institution *vis-à-vis* the executive, and to do so using a latent
variable aggregation that avoids the arbitrary weighting of linear
additive indices.[^additive] The aggregation method is largely similar
to the horizontal accountability index, *v2x_horacc*, created by
@Luhrmann2017 and the presidentialism index, *v2xnp_pres*, created by
@Sigman2018 already available in the V-Dem dataset. By creating a
separate index, however, I am able to adjust the list of included
manifest variables as well as embed the factor analysis within the
larger regression model such that the uncertainty from the latent
factor propagates as a result of the estimation process.

[^additive]: A theoretical argument could be made that the oversight
    offered by each institution is a necessary condition and that any
    index should therefore be additive with equal weighting. However,
    this would be a normative measure capturing the degree to which a
    country-year fulfills an ideal concept, rather than measuring how
    constrained an executive actually is in practice.

For the legislature and the judiciary, the two principal government
branches most able to check the power of the executive, there needs to
be both a *capacity* and *willingness* for oversight. Thus, I include
the following legislative variables: *v2lgoppart*, opposition parties
able to exercise oversight; *v2lgfunds*, legislature controls its own
resources; *v2lginvstp*, legislature investigates in practice;
*v2lgqstexp*, legislature questions executive branch officials. For
the judicial component: *v2juhcind*, high court independence;
*v2juncind*, lower court independence; *v2juhccomp*, the degree to
which the government complies with high court decisions; *v2jucomp*,
the degree to which the government complies with lower court
decisions. And for the final dimension, I include the variable
*v2exrescon*, executive respects constitution.[^lgotovst]

[^lgotovst]: Both *v2x_horacc* and *v2xnp_pres* also include the
    variable *v2lgotovst*, executive oversight by bodies other than
    the legislature such as comptroller general, general prosecutor,
    or ombudsman. While this is certainly a relevant variable, there is
    an error in V-Dem's v9.0 dataset whereby it has been cleaned
    according to when the legislature is missing. The result is that
    1,619 country-years, or approximately $13.5\%$ of observations,
    are missing. Since this pattern of missingness is not random, and
    I lack additional information predicting the values for this
    variable, I cannot implement a general imputation
    strategy. Instead, I avoid including it in the index. Due to the
    already large number of manifest variables and *v2lgotovst* being
    fairly colinear with *v2lginvstp*, legislature investigates in
    practice, with a correlation of 0.84, the result of dropping
    *v2lgotovst* is largely inconsequential.

(ref:fa-sum) Descriptive statistics for manifest variables

```{r fa-sum}
manifest <- merged.df %>%
    select(v2lgoppart, v2lgfunds, v2lginvstp, v2lgqstexp, v2juhcind,
           v2juncind, v2juhccomp, v2jucomp, v2exrescon)

q <- summary_stats(manifest) %>%
    mutate(Variable = paste0("\\textit{", Variable, "}"))

kable(q, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "(ref:fa-sum)") %>%
    kable_styling(latex_options = "striped")

```

Table \ref{tab:fa-sum} provides summary statistics for the manifest
variables. All of the variables are drawn from expert coded surveys
and have been aggregated by V-Dem using IRT models. Thus, they are all
on an unbounded scale with observations roughly falling between -3
and 3. Furthermore, the scale directions are identical and follow the
general direction of the survey answer categories for lower level
V-Dem indicators --- namely, low to high, *i.e.*, 0 is the "worst" and
the highest value the "best." For example, for *v2lginvstp*,
legislature investigates in practice, the answer category 0 denotes
that it is extremely unlikely that the legislature conducts an
investigation where the outcome would be unfavorable for the sitting
executive and the highest category, 4, denotes that it is certain or
nearly certain. On the aggregated, unbounded scale, after running the
IRT model, this translates to higher values for the point estimates
representing country-years where the legislature is more likely to
investigate.  Finally, figure \ref{fig:cor} also provides a
correlation matrix demonstrating a fairly high level of correlation
across all variables indicating some commonality in the covariation.

(ref:cor) Correlation matrix of manifest variables

```{r fig.height = 5, fig.width = 7, fig.cap = "\\label{fig:cor} (ref:cor)"}
m <- cor(manifest, use = "pairwise.complete.obs") %>%
    signif(3)

corrplot(m, "number", "upper", diag = F, tl.col = "black")

```

Given the listed manifest variables, the estimation technique is then
as follows: for $m$ country-years indexed by $i = 1 \dotsc m$ and $n$
manifest variables indexed by $j = 1 \dotsc n$, let $Z \in
\mathbb{R}^{m \times n}$ be the matrix of manifest variables measured
with an observed degree of error, $U \in \mathbb{R}^{m \times n}$, and
$Z^\ast \in \mathbb{R}^{m \times n}$ the matrix of unobserved true
values. Let $\theta$ then be the unidimensional latent factor of
length $m$ measuring executive constraints.

I first model the measurement error in the manifest variables by
assuming that the observed values are drawn from a normal distribution
where the location parameters are the unobserved true values and the
scale the observed posterior standard deviations derived from the IRT
posterior distributions. [^err]
\[ Z_{ij} \sim \mathsf{Normal}(Z^\ast_{ij}, U_{ij}) \]

[^err]: The ideal approach would be to directly embed the individual
    IRT models aggregating the manifest variables into the larger
    executive constraints measurement model [@Marquardt2019]. However,
    this prohibitively increases the run time of the model.

The latent values for the manifest variables are then modeled as
functions of the latent factor $\theta$ with a standard factor
analytic mapping function.
\[ Z^\ast_{ij} \sim \mathsf{Normal}(\gamma_j + \lambda_j \theta_i, \psi_j) \]

In the preceding equation, $\lambda$ is the variable specific factor
loadings, $\gamma$ the population intercepts, and $\psi$ the scale of
the error term. I assume local item independence such that the error
terms, $(\psi_1, \dotsc, \psi_n)$, are uncorrelated. Furthermore, for
the purposes of identifying the likelihood, the scale of the latent
factor, $\theta$, is fixed to mean $0$ and standard deviation $1$
while the factor loadings, $(\lambda_1, \dotsc, \lambda_n)$, are
constrained to be positive [@Jackman2009]. The positive constraints
follow theory since, as stated previously, all of the manifest
variables exhibit the same scale direction and should be positively
correlated with the latent factor. This means that larger values in
the point estimates for $\theta$ represent a more constrained
executive.

Similar to the regression coefficients, I select the following weakly
informative priors:
\begin{align*}
    \gamma & \sim \mathsf{Normal}(0, 5) \\
    \lambda & \sim \mathsf{LogNormal}(0, 1) \\
    \psi & \sim \mathsf{Weibull}(2, 1)
\end{align*}

In the case of the legislative variables there is the added issue of
missing observations for country-years lacking a
legislature.[^missing] For those observations, given the centrality of
the legislature as a potential constraint against the executive, the
latent factor should ideally be lower conditional on the other
manifest variables. However, since the manifest variables are
themselves unbounded latent variables, minimal values cannot be
imputed without a nonlinear transformation to a bounded scale. I thus
take a cue from *v2x_horacc* and include into the model a dichotomized
version of *v2lgbicam*, measuring how many chambers the legislature
contains, denoted by the Bernoulli distributed variable, $w$, where
country-years with a sitting legislature are given a $1$ and
country-years where the legislature is missing or has been suspended,
are assigned $0$. The probability of a legislature being present,
$\pi_i$, is then a linear function of the latent factor, $\theta$,
transformed by the inverse logit such that $\pi_i \in [0, 1]$. The
practical effect is that where there is data for the legislative
manifest variables, the latent factor is informed by the measured
institutional quality of the legislature. Where data is missing due to
the suspension or lack of a legislature, the latent factor is down
weighted by the linear function predicting $w$, conditional on the
observed values of the remaining manifest variables. The resulting
model specification, including priors, can be summarised as follows:
\begin{align*} w_i & \sim \mathsf{Bernoulli}(\pi_i) \\ \pi_i & =
\mathsf{logit}^{-1}(\eta + \delta \theta_i) \\ \eta & \sim
\mathsf{Normal}(0, 5) \\ \delta & \sim \mathsf{LogNormal}(0, 1)
\end{align*}

[^missing]: Missingness due to the lack of of a particular
    institution, for example the suspension of a high court, is an
    issue for the other manifest variables as well. However, only the
    existence of the legislature is coded by V-Dem as a separate
    variable. Furthermore, practically speaking, none of the other
    manifest variables exhibit the same systematic pattern of
    missingness in the dataset as the legislative variables. Whereas
    there are 1608 country-years lacking a legislature, or
    approximately $13.4\%$ of all post-1945 country-years, only 115
    are missing data for at least one of the judicial variables
    ($1\%$) and only 58 are missing for *v2exrescon* ($0.5\%$). I do
    not implement any imputation strategy for the manifest
    variables. The rationale is that, disregarding data processing
    issues, missingness in V-Dem can arise either from a lack of
    expert coders or lack of information. The latter scenario will
    most likely be correlated with country-years where there is a
    crisis or significant political instability, such as coups or
    civil war, where extreme values are more probable. Since it cannot
    be assumed that the values are completely missing at random, the
    data generative process for missingness cannot be modeled without
    potentially introducing bias into the estimates.

### Additional Covariates

As mentioned during the theoretical discussion of possible mechanisms,
the presence of executive constraints alone are insufficient in
determining the outbreak of conflict. One possible moderator is the
relative capacity of the state. For the present analysis I follow
@Bartusevic2018 in using the Composite Indicator of National
Capability (CINC) from the National Material Capabilities v5.0 dataset
(1816-2012) [@Singer1987]. The CINC index is aggregated from several
demographic, industrial, and military indicators in order to capture
the available resources a country can bring to bear in the event of a
conflict. This is an imperfect measure that largely misses the state's
bureaucratic and administrative capacity [@Hendrix2010]. [^wbgi]
However, given that the dimension of interest for this analysis is the
repressive capacity of the state, material capabilities, especially
military resources, should still provide an adequate proxy.

[^wbgi]: Furthermore, a common issue with many measures of
    bureaucratic quality is missingness.  For example, the World Bank
    governance indicator on government effectiveness extends back only
    until 1996 [@Kaufmann2010]. Given that measures of institutional
    quality often change fairly slowly, punctuated only by abrupt
    events such as coups or revolutions, an extended time series is
    desirable. Furthermore, in the specific case of bureaucratic
    quality, there is the added issue of endogeneity. An unconstrained
    executive may seek to promote government corruption and undermine
    both meritocratic recruitment into the bureaucracy as well as its
    impartiality.

To capture the relative ease of mobilizing fighters, especially in
terms of being able to offer economic selective benefits, I include
GDP per capita. Specifically, I use the natural log of the $RGDP^e$
variable, constant price real GDP, from Penn World Tables (PWT) v9.1
(1950-2017) to allow for both cross-country and cross-temporal
comparisons of GDP per capita [@Feenstra2015].[^gdp] The justification
for taking the natural log is that the marginal effect of the variable
should decrease with higher values. In other words, the difference
between the effects of relatively high levels of GDP per capita should
be significantly smaller than the difference between the effects of
low levels of GDP per capita. Finally, I also include an interaction
term between GDP per capita and the latent measure for executive
constraints to capture any potential conditional relationship to
specifically test whether the effect from executive constraints
diminishes with lower levels of GDP per capita.

[^gdp]: Prior to taking the natural log, the GDP measure is
    transformed to GDP per capita by dividing by the total population,
    which is also available in PWT.  Note, PWT has two real GDP
    measures available: $RGDP^o$, or output side real GDP, and
    $RGDP^e$, or expenditure side real GDP. The former roughly
    captures a country's productive capacity, while the latter is
    better suited to measuring living standards [@Feenstra2015]. I use
    $RGDP^e$ since I would argue that it serves as a better proxy for
    the potential recruitment costs for rebel leaders.

In addition, I also include a number of standard covariates drawn
largely from previous literature. The list of indicators are not meant
to provide an all encompassing theory of civil conflict, but rather
roughly capture the opportunity structures for rebellion.  This
includes a single year lagged measure of GDP per capita growth from
PWT in order to capture short term economic shocks, which may affect
both the strength and legitimacy of the state [@Bazzi2014]. I also
include a measure of mean elevation from the CShapes dataset
[@Weidmann2010] as made available through the GROWup data platform
[@Girardin2015] to capture terrain advantageous for conducting an
insurgency [@Fearon2003]. In addition, I employ a measure of
population density derived from population data from PWT and the area
in square kilometers for each country from CShapes. Countries with a
higher population density may require a more extensive and competent
central administration and police force in order to suppress violent
dissent [@Fearon2003]. There is also the possibility that a high
population density increases the competition for resources. If the
state is identified as the source of a particular scarcity, the risk
of civil conflict may increase. Similar to GDP per capita, I also
transform mean elevation and population density by the natural log
with the expectation that the scale increase should diminish as values
get larger. Finally, I include measures of ethnic fractionalization in
the form of the number of politically relevant ethnic groups and the
percent of the population facing discrimination as taken from the
Ethnic Power Relations dataset [@Vogt2015].[^epr] Where there are
significant ethnic cleavages and state sponsored discrimination,
ethnic ties can facilitate collective action thereby lowering the
mobilization costs of rebel groups [@Cederman2011].

[^epr]: In this context, an ethnic group is considered politically
    relevant if "either at least one significant political actor
    claims to represent the interests of that group in the national
    political arena or if group members are systematically and
    intentionally discriminated against in the domain of public
    politics" while discrimination denotes the formal political
    exclusion of a specific ethnic group with citizen rights
    [@EPR2019].

As mentioned previously, to control for time dependence --- past and
current civil conflicts can impact the present risk of conflict onset
as well as the current level of executive constraints -- I include the
number of peace years since the last conflict incident.[^time] I also
include a dummy variable for ongoing high intensity conflict since
devastating civil wars may have a particularly destabilizing
effect.[^intensity] To account for spillover effects, I include a
dummy variable for whether a neighbouring country is currently
experiencing a civil conflict.[^geo] Finally, one notable lack is a
general measure of political instability, which I omit since it could
be highly correlated with executive constraints and would potentially
bias results, at least for years of crisis where the risk of civil
conflict is highest. I do, however, include a dummy variable for the
independence year since newly sovereign states may be especially
vulnerable.

[^time]: To be more specific, civil conflicts can be highly
    destabilizing and severely affect the governing and military
    capacity of a state, increasing the risk for an additional
    conflict onset. Moreover, in times of particularly intense
    conflict, wartime powers may be granted to the executive bypassing
    normal institutional constraints. Thus, ongoing conflict is a
    potential confounder since it can influence both the probability
    of the outcome as well as the executive constraints measure. For
    this purpose, the peace years variable is calculated as the number
    of years since the last conflict-year, including conflict-years
    from extrasystemic conflicts. For countries left censored due to
    the dataset starting date, I set the start of the peace year
    calculation to 1946, the end of World War II. The rationale is
    that World War II was a significant international event that
    reshaped the world order and had domestic consequences for even
    countries not directly involved. For newly independent countries,
    I use the independence date as the start year.

[^intensity]: High intensity conflict meaning conflict-years with at
    least 1000 battle deaths.

[^geo]: Country neighbours for each year were determined according to
    CShape. Pre-independence conflicts were considered as neighbouring
    to the countries bordering the territory location of the conflict,
    not the government location.

Table \ref{tab:reg-stats} lists the full descriptive statistics for
all regression covariates.

(ref:reg-stats) Descriptive statistics for regression covariates

```{r reg-stats}
df <- merged.df %>%
    select(rgdpepc, rgdpepc_gro, cinc, discrimpop, pop_density,
           meanelev, rlvt_groups_count, neighbour_conflict,
           major_ongoing, independence)

q <- summary_stats(df)

kable(q, "latex", align = c("l", "r", "r", "r", "r", "r"),
      booktabs = T, linesep = "", escape = T,
      caption = "(ref:reg-stats)") %>%
    kable_styling(latex_options = "striped")

```

## Results

The analysis of the results is split into several sections. First, the
parameter estimates from the executive constraints measurement model
will be presented. This will then be followed by the parameters
pertaining to the logistic regression before concluding with an
assessment of model fit and the presentation of several alternative
specifications for the dependent variable, conflict onset.

### Measurement Model

(ref:fpars) Executive constraints measurement model parameter
    estimates with 80 percent posterior credible intervals.

```{r fig.height = 4, fig.cap = "\\label{fig:fpars} (ref:fpars)"}
load("./posteriors/full_model/input.RData")
fa <- read_post("./posteriors/full_model/fa_posteriors.csv")

lg <- colnames(lg_mm$obs)
nonlg <- colnames(nonlg_mm$obs)

lambda <- extract(fa, pars = "lambda")
colnames(lambda) <- c(lg, nonlg)

psi <- extract(fa, pars = "psi")
colnames(psi) <- c(lg, nonlg)

gamma <- extract(fa, pars = "gamma")
colnames(gamma) <- c(lg, nonlg)

delta_eta <- extract(fa, pars = c("delta", "kappa"))

p1 <- plot_pars(lambda, "") + ggtitle(expression(lambda))
p2 <- plot_pars(psi, "") + ggtitle(expression(psi))
p3 <- plot_pars(gamma, "") + ggtitle(expression(gamma))
p4 <- plot_pars(delta_eta, "") + ggtitle(expression(delta * "-" * kappa)) +
    theme(axis.text.x = element_text(angle = 0)) +
    scale_x_discrete(labels = c(expression(delta), expression(kappa)))

grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

Table \ref{tab:fa} and figure \ref{fig:fpars} summarise the posterior
draws for the measurement model parameters with 80 percent posterior
quantile intervals. According to the results, the factor loadings,
$\lambda$, for each manifest variable are fairly comparable, with the
exception of the lower value for *v2lgfunds* ("legislature controls
its own resources") and to a lesser extent *v2lgqstexp* ("legislature
questions executive branch officials in practice"). Relative to the
other manifest variables, both items are less informative in inferring
the values of the latent factor --- a $\lambda$ of zero would indicate
that the latent factor does not explain any of the variation in the
manifest variable. As for the variable specific intercepts, $\gamma$,
they are largely clustered around zero as to be expected given that
the manifest variables were standardized prior to the model run. Where
deviations do occur, most notably with several of the legislative
variables, it is because of the measurement error component of the
model shifting the mean for a specific variable away from zero. For
example, after adjusting for measurement error, the mean for
$v2lgfunds$ is on average slightly higher than the remaining manifest
variables, hence the negative intercept.

(ref:fa) Executive constraints measurement model parameter
    estimates with 80 percent posterior credible intervals.

```{r fa, fig.height = 6, fig.width = 6}
fa_pars <- extract(fa, c("lambda", "gamma", "psi", "delta", "kappa")) %>%
    post_summarise(probs = c(0.10, 0.9)) %>%
    mutate(parameter = sub("\\[\\d+\\]", "", parameter) %>% sprintf("$\\%s$", .),
           interval = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3)),
           manifest = c(rep(c(lg, nonlg), 3),
                        rep("v2lgbicam", 2))) %>%
    select(-`10%`, -`90%`) %>%
    pivot_wider(id_cols = manifest, names_from = parameter, values_from = interval)

# I absolutely hate tibbles
fa.mm <- select(fa_pars, -manifest) %>% as.matrix
rownames(fa.mm) <-sprintf("\\textit{%s}", fa_pars$manifest)

kable(fa.mm, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "(ref:fa)") %>%
    kable_styling(latex_options = c("striped", "scale_down"), full_width = F,
                  protect_latex = F)
```

The $\delta$ and $\kappa$ parameters correspond to the slope and
intercept in the linear predictor for the presence of a legislature,
*v2lgbicam*. The effect of their inclusion can be seen in figure
\ref{fig:bicam}. The latent factor, $\theta$, follows the
directionality of the manifest variables, greater values indicate a
more constrained executive and lower values vice versa. The density
plots indicate that on average, for country-years where there is no
legislature, *i.e.* $v2lgbicam$ is zero, the latent factor is lower,
although still conditional on the values for the remaining manifest
variables.

(ref:bicam) Posterior density plots for the executive constraints
    measure, $\theta$.

```{r fig.width = 6, fig.height = 2, fig.cap = "\\label{fig:bicam} (ref:bicam)"}
theta <- extract(fa, "theta")

groups <- ifelse(lgbicam == 1, "Legislature Present", "No Legislature")
plot_dens(theta, groups = groups)

```

The final portion of the measurement model is the measurement error
adjustment. As described in the research design section, in order to
incorporate the posterior uncertainty from the V-Dem IRT aggregation,
the observed values for the manifest variables, $Z$, were assumed to
be drawn from normal distributions where the means were the unobserved
true values, $Z^{\ast}$, and the standard deviations the observed
posterior standard deviations as reported by V-Dem. The practical
effect is that for a given observation where the reported error is
larger, the uncertainty in the $Z^{\ast}_{ij}$ estimate increases and
the mean is drawn towards the regression line in the factor
analysis. In other words, the model is more skeptical of outliers in
the manifest variables. This posterior shrinkage can be seen in figure
\ref{fig:fa-line}. Figure \ref{fig:err} meanwhile plots the range of
the upper and lower bounds of the $80$ percent posterior quantile
intervals for an example manifest variable in $Z^{\ast}$ against the
observed V-Dem standard deviations as well as the difference between
the estimated posterior mean and the observed mean in order to further
demonstrate this effect.

(ref:fa-line) Posterior shrinkage for *v2juncind*. The observed point
    estimates are the data points reported by V-Dem.

```{r fig.cap = "\\label{fig:fa-line} (ref:fa-line)"}
err <- read_post("posteriors/full_model/err_posteriors.csv")

target <- colnames(nonlg_mm$obs)[1]
nonlg_est <- extract(err, "nonlg_est") %>%
    post_summarise %>%
    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
    mutate(interval = codehigh - codelow,
           se = nonlg_mm$se[, 1],
           obs = nonlg_mm$obs[, 1],
           diff = obs - median)

idx <- sample(1:ncol(theta), 50)
theta_medians <- theta[, idx] %>% apply(2, median)

s <- seq(min(theta_medians) - 0.1, max(theta_medians) + 0.1, 0.01)

x <- data.frame(point = c(nonlg_est$obs[idx], nonlg_est$median[idx]),
                group = rep(1:length(idx), 2),
                factor = c(rep("Observed Point Estimate", length(idx)),
                           rep("Estimated Posterior Median", length(idx))) %>% as.factor,
                theta = rep(theta_medians, 2))

y <- data.frame(yhat = median(gamma[, 6]) + median(lambda[, 6]) * s,
                theta = s)

ggplot(x, aes(theta, point, group = group, col = factor)) +
    geom_line(alpha = 0.3) +
    geom_point(alpha = 0.8) +
    geom_line(data = y, aes(theta, yhat, group = NA), colour = "black", alpha = 1) +
    theme_tufte() +
    scale_color_colorblind() +
    theme(legend.title = element_blank()) +
    ylab(target) +
    xlab("Theta")

```

(ref:err) On the left, the posterior uncertainty for *v2juncind* as
    estimated by the model plotted against the observed error reported
    by V-Dem as *v2juncind_sd*. On the right, the difference between
    the point estimates estimated by the model and the point estimates
    reported by V-Dem versus the observed error for *v2juncind*.

```{r fig.cap = "\\label{fig:err} (ref:err)"}
p1 <- ggplot(nonlg_est, aes(se, interval)) +
    geom_point(col = "#56B4E9", alpha = 0.2, shape = 1) +
    theme_tufte() +
    ylab("Posterior CI Range") +
    xlab("Observed Error")

p2 <- ggplot(nonlg_est, aes(se, diff)) +
    geom_point(col = "#56B4E9", alpha = 0.2) +
    theme_tufte() +
    ylab("Pointwise difference") +
    xlab("Observed Error")

grid.arrange(p1, p2, ncol = 2)

```

#### Executive Constraints Estimates

In total there are `r nrow(final.df) %>% prettyNum(big.mark = ",")`
country-year estimates of $\theta$, comprised of
`r n_distinct(final.df$country_name)` unique countries and scaled from
approximately $-2$ to $2$. Since the complete time series for every
single country can not be easily visualised, an exploratory overview
of the performance of the executive constraints measure is provided in
the form of several different plots.  The first, figure
\ref{fig:theta}, is a restricted time series plot consisting of only
China, Brazil, Nigeria, Norway, and Russia. The countries were chosen
to represent a diverse set of geographic regions as well as political
contexts. From the plot, China, being a single party authoritarian
state, is consistently ranked the worst in executive constraints. The
decline for Russia largely follows the ascent of Vladimir Putin to the
presidency at the end of 1999 and his gradual consolidation of power
in the decades that followed.[^russia] For both Nigeria and Brazil,
the measure is able to distinguish between democratic rule and periods
of military dictatorship such as the era of military rule in Brazil
following the 1964 coup d'etat and the period following Johnson
Aguiyi-Ironsi's seizure of power in the 1966 coup d'etat in
Nigeria. The last country, Norway is an example of a consolidated
polyarchy for the time period under consideration with consistently
strong constraints against the executive office which are accurately
captured by the measure.

[^russia]: As can be seen in the plot, the observations for Russia
    begin in 1991 with the fall of the Soviet Union. There is data
    available in the V-Dem dataset to extend back the executive
    constraints measure to cover the years of the Russian Soviet
    Federative Socialist Republic. However, those observations are
    lost when merging all of the covariates into the final dataset due
    to the lack of data in PWT, most likely stemming from the
    difficulty of converting the Soviet measure of economic
    performance, Net Material Product, to GDP.

(ref:theta) Executive constraints over time for several selected
    countries.

```{r fig.width = 6, fig.height = 4, fig.cap = "\\label{fig:theta} (ref:theta)"}
theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
    rename(codelow = `10%`, median = `50%`, codehigh = `90%`)

theta_summarised <- theta_summarised[final.df$reduced_idx, ] %>%
    mutate(country_name = final.df$country_name,
           gwid = final.df$gwid,
           year = final.df$year)

targets <- c("Brazil", "China", "Nigeria", "Russia", "Norway")

filter(theta_summarised, country_name %in% targets) %>%
    ggplot(aes(year, median, fill = country_name)) +
    theme_tufte() +
    scale_fill_colorblind() +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
    theme(legend.title = element_blank())

```

For a full cross-country comparison, figure \ref{fig:map} provides a
snapshot of the posterior medians for $\theta$ for all countries in
the year 2012. Lighter colors represent countries with higher levels
of executive constraints, darker colors vice versa, and countries
which are grey are missing from the dataset. Given the plot, regional
differences are readily apparent; for example, many of the former
Soviet Socialist Republics, with the notable exception of the Baltic
states, are markedly lower, *i.e.* lacking in effective constraints
against the executive, than the former Eastern bloc countries. This
plot does not depict, however, the uncertainty in the estimates.

(ref:map) Executive constraints in 2012 for all countries present in
    the dataset. Missingness indicated by the color grey.

```{r results = "hide", fig.width = 8, fig.cap = "\\label{fig:map} (ref:map)"}
map_data <- st_read("data/raw/cshapes_0.6/cshapes.shp") %>%
    filter(GWEYEAR == max(GWEYEAR)) %>%
    select(country_name = CNTRY_NAME, gwid = GWCODE)

df <- filter(theta_summarised, year == max(year)) %>%
    select(-country_name) %>%
    right_join(map_data) %>%
    st_as_sf

ggplot(df) +
    geom_sf(aes(fill = median)) +
    theme_tufte() +
    theme(legend.title = element_blank())

```

```{r results = "hide"}
chile_idx <- filter(final.df, country_name == "Chile", year %in% c(2000, 2010)) %$%
    reduced_idx

chile_p <- signif(mean(theta[, chile_idx[1]] < theta[, chile_idx[2]]) * 100, 2)

tunisia_idx <- filter(final.df, country_name == "Tunisia", year %in% c(2000, 2010)) %$%
    reduced_idx

tunisia_p <- signif(mean(theta[, tunisia_idx[1]] < theta[, tunisia_idx[2]]) * 100, 2)
```

Finally, to expand upon the country selection from figure
\ref{fig:theta} and provide a temporal comparison, figure
\ref{fig:tie} plots the $\theta$ estimates for 2010 versus 2000 by
country. A full list of all countries is available in [Appendix
C](#appendix-c). Note, the statistical significance of differences
between years for a single country cannot strictly be judged by
overlapping credible intervals [@Schenker2001]. For example, Chile,
despite having overlapping credible intervals, still has a posterior
probability of approximately $`r chile_p`$ percent that the estimate
for 2000 is lower than 2010. In the case of Tunisia with even more
visual overlap, there is a $`r tunisia_p`$ percent posterior
probability that the estimate for 2010 is higher than 2000.

(ref:tie) Executive constraints in 2000 versus 2010 for several
    selected countries.

```{r fig.cap = "\\label{fig:tie} (ref:tie)"}
targets <- c("Algeria", "Argentina", "Armenia", "Chile", "Colombia",
             "Egypt", "Georgia", "Ghana", "Hungary", "Indonesia",
             "Iran", "Israel", "Japan", "Kenya", "Mexico", "Poland",
             "Portugal", "Romania", "Rwanda", "Turkey", "Tunisia",
             "United States of America", "Zimbabwe")

df <- filter(theta_summarised, year %in% c(2010, 2000),
             country_name %in% targets) %>%
    mutate(year = as.factor(year),
           country_name = ifelse(country_name == "United States of America",
                                 "USA", country_name))

tfplot(df) + xlab("Theta")
```

### Onset Regression

Figure \ref{fig:reg} plots the slope coefficients from the logistic
regression linking the latent construct for executive constraints to
the outcome, civil conflict onset, on the log-odds scale. Credible
intervals are summarised in the same way as previous plots with 80
percent quantile intervals, and exact values are reported in table
\ref{tab:reg} together with the posterior probabilities that each
$\beta$ estimate is greater than zero and less than zero. Recall that
the variables are standardized to have mean 0 and standard deviation
1, with the exception of the ordinal/binary variables Relevant Group
Count, Neighbouring Conflict, Ongoing Major Conflict, and
Independence.

(ref:reg) Regression slopes with 80 percent posterior credible
    intervals and the posterior probability of each parameter being
    less than and greater than zero.

```{r reg}
reg <- read_post("posteriors/full_model/reg_posteriors.csv")

beta <- extract(reg, "beta")
colnames(beta) <- c("exec_constraints", colnames(X),
                    sprintf("exec * %s", colnames(X)[interaction_idx]))

# Probabilities that Beta < 0 and Beta > 0
probs <- apply(beta, 2, function(x) {
    list(below = mean(x < 0) %>% round(3) %>% sprintf("%.3f", .),
         above = mean(x > 0) %>% round(3) %>% sprintf("%.3f", .))
})

beta_s <- post_summarise(beta, probs = c(0.1, 0.9)) %>%
    mutate(`Interval Estimate` = sprintf("[%.3f, %.3f]",
                                         signif(`10%`, 3), signif(`90%`, 3)),
           `$Pr(\\beta < 0)$` = lapply(probs, `[[`, "below") %>% unlist,
           `$Pr(\\beta > 0)$` = lapply(probs, `[[`, "above") %>% unlist) %>%
    select(-`10%`, -`90%`, -parameter)

rownames(beta_s) <- colnames(beta) %>% gsub("_", "\\\\_", .)

kable(beta_s, "latex", align = c("c", "r", "r"), booktabs = T,
      linesep = "", escape = F, caption = "(ref:reg)") %>%
    kable_styling(latex_options = "striped", "scale_down")
```

(ref:fig-reg) Regression slopes with 80 percent posterior credible
    intervals.

```{r fig.cap = "\\label{fig:reg} (ref:fig-reg)"}
plot_pars(beta, hline = T)
```

For the $\beta$ parameter for executive constraints there is a `r
as.numeric(beta_s[1, 2]) * 100` percent posterior probability that the
parameter is less than zero and a 80 percent posterior probability
that an increase of one standard deviation decreases the log-odds of
civil conflict onset between $`r beta_s[1, 1]`$. A one standard
deviation increase in this context corresponds roughly to the
difference between the level of executive constraints for Zambia in
1967 under the administration of Kenneth Kaunda right before the
banning of opposition parties following the 1968 election, with a
score of approximately 0, and Greece in 1976 shortly after the end of
military rule and the establishment of a democratic republic with a
score of approximately 1.

```{r results = "hide"}
argentina <- filter(merged.df, country_name == "Argentina", year == 1989) %$%
    prettyNum(rgdpepc, big.mark = ",")

slovakia <- filter(merged.df, country_name == "Slovakia", year == 2003) %$%
    prettyNum(rgdpepc, big.mark = ",")

cmb <- filter(merged.df, country_name == "Cambodia", year == 1987) %>%
    select(pop, area_sqkm) %>%
    mutate(area_sqkm = signif(area_sqkm, 3),
           pop = signif(pop, 3))

india <- filter(merged.df, country_name == "India", year == 1976) %>%
    select(pop, area_sqkm, meanelev) %>%
    mutate(pop = signif(pop, 3),
           area_sqkm = signif(area_sqkm, 3),
           meanelev = signif(area_sqkm, 3))

georgia <- filter(merged.df, country_name == "Georgia", year == 2012) %$%
    signif(meanelev, 3)

russia <- filter(merged.df, country_name == "Russia", year == 2012) %$%
    signif(meanelev, 3)

thai <- filter(merged.df, country_name == "Thailand", year == 2002) %$%
    { discrimpop * 100 }

bhutan <- filter(merged.df, country_name == "Bhutan", year == 1989) %$%
    { discrimpop * 100 }

```

The slopes for the remaining covariates generally follow the intuitive
expectations of their relationship to civil conflict onset with
varying degrees of certainty. For example, GDP per capita is negative
to which a one unit increase from 0 to 1 is roughly similar to the
difference between Argentina in 1989 at the height of a hyperinflation
crisis with a value of $`r argentina`$ 2011 US dollars and Slovakia in
2003 with a value of $`r slovakia`$ 2011 US dollars at the beginning
of a period of high GDP growth. In other words, an increase in GDP per
capita on the logarithmic scale decreases the log-odds of conflict
onset. GDP per capita growth and CINC are also negative, but the
estimates are significantly less reliable. The posterior probability
of being less than zero is $`r beta_s[3, 2]`$ for the former and only
$`r beta_s[4, 2]`$ for the latter.

Population density, mean elevation, number of politically relevant
ethnic groups, proportion of the population facing state
discrimination, presence of at least one neighbouring conflict, and
the dummy variable for independence year all increase the risk of
civil conflict onset. A one unit increase in population density is
similar to going from Cambodia in 1987 with a population of $`r
cmb$pop`$ million and an area of $`r cmb$area_sqkm`$ square kilometers
to India in 1976 with a population of $`r india$pop`$ million and an
area of $`r india$area_sqkm`$ square kilometers. A one unit increase
for mean elevation roughly matches the difference between Russia at
$`r russia`$ meters and Georgia at $`r georgia`$ meters. And, a one
unit increase for proportion of the population facing discrimination
is roughly equivalent to Thailand from 2002 until present with $`r
thai`$ percent and Bhutan from 1989 until present with $`r bhutan`$
percent.

Finally, the GDP per capita and executive constraints interaction term
and the dummy variable for ongoing high intensity conflict decrease
the risk of conflict onset. This latter result, although surprising,
may be due to the subsuming nature of high intensity civil wars. In
the face of an all-encompassing civil war, there may be a plethora of
active factions with divergent interests; however, their political
goals and interests are structured according to the current
incompatibility between the dominant political actors. Thus, there may
be new armed groups emerging who are willing to challenge to the
state, but that emergence may not necessarily be coded as a new civil
conflict. It should be noted, however, that there is a high degree of
uncertainty in the estimate with the 80 percent posterior credible
interval ranging from -0.558 to 0.064.

#### Random Effects

There are three sets of random effects included in the regression
model. The first to be presented in figure \ref{fig:year} is the year
intercepts with 80 percent posterior credible intervals. There is no
discernible time trend in the year specific variability. The estimates
are strongly pulled towards zero as a result of the partial pooling
with upticks, such as the initial years of the 1990s during the
dissolution of the Soviet Union and Yugoslavia, roughly mirroring the
global incidence rate of onsets.

(ref:year) Log-odds year intercepts with 80 percent credible intervals

```{r fig.cap = "\\label{fig:year} (ref:year)"}
Z_year <- extract(reg, "Z_year")
colnames(Z_year) <- { unique(final.df$year) + 1 } %>% sort

breaks <- seq(min(final.df$year), max(final.df$year), by = 4)
plot_pars(Z_year, hline = T) +
    scale_x_discrete(breaks = breaks)

```

```{r results = "hide"}
alpha <- extract(reg, "alpha")

alpha_s <- quantile(alpha, probs = c(0.1, 0.9)) %>% signif(3)
alpha_interval <- sprintf("[%g,%g]", alpha_s[1], alpha_s[2])
```

Figure \ref{fig:country} presents the country intercepts. Since all of
the countries cannot be easily visualized at once only the top ten and
bottom ten are depicted. From the results it is clear that there is
considerably less uncertainty in the estimates for the top
countries. The intercepts for these countries are consistently above
zero meaning that the risk of civil conflict onset is higher than that
predicted solely by the other predictors. Unsurprisingly, these
countries tend to experience the highest number of civil conflict
onsets in the dataset which, conditional on the data, accounts for the
higher baseline risk. The countries at the bottom have experienced
relatively few, if any, conflict onsets relative to their values for
the other predictors. The majority of intercepts for these countries
overlap zero possibly because the global intercept is already fairly
low, `r alpha_interval` (80 percent log-odds CI).

(ref:country) Log-odds country intercepts with 80 percent credible
    intervals

```{r, fig.cap = "\\label{fig:country} (ref:country)"}

Z_country <- extract(reg, "Z_country")
colnames(Z_country) <- unique(final.df$country_name) %>% sort

x <- apply(Z_country, 2, median) %>% order(decreasing = F)
top <- head(x, 10)
bottom <- tail(x, 10)

fn <- function(x, title) {
    plot_pars(x, hline = T) +
        coord_flip() +
        theme(axis.title.y = element_blank()) +
        ggtitle(title)
}

p1 <- fn(Z_country[, top], "Bottom 10")
p2 <- fn(Z_country[, bottom], "Top 10")

grid.arrange(p1, p2, ncol = 2)

```

```{r results = "hide"}
gp <- extract(reg, c("rho", "eta")) %>% post_summarise(probs = c(0.1, 0.9)) %>%
    mutate(interval = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3)))
```

Finally, figure \ref{fig:peace} plots the peace year intercepts.[^gp]
The strongest effect is found in the years immediately following an
occurrence of intrastate political violence resulting in an increased
risk for an additional onset.  Approximately a decade after the last
incidence of violence, the effect is drawn towards zero. The
additional spike around approximately 30 years follows an uptick in
conflict onsets in the dataset mostly among authoritarian regimes such
as, for example, the 1989 coup d'etat overthrowing the Stroessner
regime in Paraguay, the Saur Revolution in 1978 in Afghanistan, and
the start of the Nepalese Civil War in 1996. After 30 years, the
effect of each additional year of peace tends to be negative; however,
there is a very high degree of uncertainty in the estimates that
undermines any general conclusion regarding a pacifying effect in the
model.

[^gp]: The 80 percent posterior credible intervals for the
    hyperparameters from the gaussian process modeling the effect of
    peace years were
    `r filter(gp, parameter == "rho") %$% interval` and
    `r filter(gp, parameter == "eta") %$% interval`
    for the length-scale ($\rho$) and the marginal standard deviation
    ($\eta$) respectively.

(ref:peace) Log-odds peace year intercepts with 80 percent credible
    intervals

```{r, fig.cap = "\\label{fig:peace} (ref:peace)"}
f <- extract(reg, "f")
colnames(f) <- unique(final.df$lpeace_yrs) %>% sort

plot_pars(f, hline = T) +
    scale_x_discrete(breaks = seq(0, ncol(f), by = 4))
```

#### Interaction Effect

```{r results = "hide"}
gdp_rwa <- filter(merged.df, country_name == "Rwanda", year == 2007) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_phl <- filter(merged.df, country_name == "Philippines", year == 2008) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_po <- filter(merged.df, country_name == "Portugal", year == 2009) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")
```

To interpret the interaction term between executive constraints and
GDP per capita I calculate the marginal effect of $\theta$, executive
constraints, on the predicted probability of civil conflict onset for
different levels of GDP per capita while holding constant the other
covariates. This means that I take the mean of GDP per capita growth,
CINC, the proportion of discriminated population, population density,
mean elevation, and the number of politically relevant ethnic groups
rounded to the nearest integer. I also set the dummy indicators for
independence, major ongoing conflict, and neighbouring conflict to
zero following the majority of observations in the dataset. I set the
random intercepts for year and country to zero, which is the average
effect. For peace years I take the intercept corresponding to the
average number of peace years rounded to the nearest integer. The
levels of GDP per capita were select based on the 10\%, 50\%, and 90
\% quantiles of the standardized log transformed values. These levels
roughly correspond to the GDP per capita of Rwanda in 2007 at
$`r gdp_rwa`$ 2011 US dollars, the Philippines in 2008 at $`r gdp_phl`$
2011 US dollars, and Portugal in 2009 at $`r gdp_po`$ 2011 US
dollars. In the dataset there are 60 conflict episode onsets below the
first level, 152 between the first and second, 55 between the second
and third, and 4 above the third. The results are plotted in figure
\ref{fig:int}.

(ref:int) Interaction effect between executive constraints and GDP per
    capita with 80 percent credible intervals.

```{r fig.cap = "\\label{fig:int} (ref:int)"}
# Interaction plot
medians <- apply(theta, 2, quantile, probs = 0.5)
lower_bound <- min(medians) %>% round(digits = 1)
upper_bound <- max(medians) %>% round(digits = 1)

theta_seq <- seq(lower_bound, upper_bound, .01)
state <- quantile(X[, "rgdpepc"], probs = c(0.1, .5, 0.9))

marginal_data.df <- as.data.frame(X) %>%
    summarise_all(~mean(.)) %>%
    mutate(rlvt_groups_count = round(rlvt_groups_count),
           neighbour_conflict = 0,
           major_ongoing = 0,
           independence = 0)

peace_yr <- f[, mean(final.df$lpeace_yrs) %>% round]

fn <- function(x, y) {
    with(marginal_data.df,
         alpha +
         beta[, 1] * x +
         beta[, 2] * y +
         beta[, 3] * rgdpepc_gro +
         beta[, 4] * cinc +
         beta[, 5] * discrimpop +
         beta[, 6] * pop_density +
         beta[, 7] * meanelev +
         beta[, 8] * rlvt_groups_count +
         beta[, 9] * neighbour_conflict +
         beta[, 10] * major_ongoing +
         beta[, 11] * independence +
         beta[, 12] * x * y +
         peace_yr)
}

df <- lapply(state, function(y) {
    mclapply(theta_seq, function(point) {
        p_hat <- boot::inv.logit(fn(point, y)) * 100
        post_summarise(p_hat, probs = c(0.1, 0.5, 0.9)) %>%
            mutate(theta = point,
                   y = y)
    }, mc.cores = 4) %>% bind_rows
}) %>% bind_rows

df %<>%
    rename(codelow = `10%`,
           median = `50%`,
           codehigh = `90%`) %>%
    mutate(y = case_when(y == state[1] ~ "Low GDP per Capita",
                         y == state[2] ~ "Median GDP per Capita",
                         y == state[3] ~ "High GDP per Capita") %>%
               factor(levels = c("Low GDP per Capita",
                                 "Median GDP per Capita",
                                 "High GDP per Capita")))

ggplot(df, aes(theta, median)) +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
    facet_grid(. ~ y) +
    theme_tufte() +
    xlab("Executive Constraints") +
    ylab("Predicted Probability")

```

From the listed plots it is clear that for medium to high income
country-years the overall effect of executive constraints when
interacted with GDP per capita on the predicted probability of
conflict onset is negative --- that is to say a more constrained
executive is associated with a decreased risk of civil conflict onset.
For example, for country-years with a relatively high level of GDP per
capita, going from the worst to best in executive constraints
decreases the predicted probability from [1.5, 4.5] down to below
0.5. This effect diminishes, however, for lower values of GDP per
capita to the point that executive constraints are practically
indeterminate to the risk of conflict onset for country-years situated
around the 10\% quantile for GDP per capita due to the high degree of
posterior uncertainty. In the plot depicting low GDP per capita, there
is, nonetheless, a slight increase in the median predicted
probability, which begs the question: for GDP per capita values lower
than the 10\% quantile is there a distinguishable reversal in the
relationship between executive constraints and civil conflict such
that there is actually a *positive* increase in the predicted
probability for higher levels of executive constraints?

(ref:gdp) Difference in predicted probabilities between the lowest and
    highest level of executive constraints for different values of GDP
    per capita.

```{r fig.cap="\\label{fig:gdp} (ref:gdp)"}
rgdpepc <- seq(min(X[, "rgdpepc"]), state[1], 0.01)
df <- mclapply(rgdpepc, function(y) {
    low <- boot::inv.logit(fn(lower_bound, y)) * 100
    high <- boot::inv.logit(fn(upper_bound, y)) * 100

    post_summarise(high - low, probs = c(0.1, 0.5, 0.9)) %>%
        mutate(y = y,
               prob = mean(high > low))
}, mc.cores = 4) %>% bind_rows

df %<>%
    rename(codelow = `10%`, median = `50%`, codehigh = `90%`) %>%
    mutate(interval = sprintf("[%s, %s]",
                              fmt(codelow, 3),
                              fmt(codehigh, 3)))

df$original <- exp(df$y * sd(final.df$rgdpepc) + mean(final.df$rgdpepc))

ggplot(df, aes(original, median)) +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
    theme_tufte() +
    xlab("GDP per Capita")
```

To investigate this possibility, \ref{fig:gdp} plots the pointwise
difference between the predicted probability of civil conflict given
the highest level of executive constraints and the predicted
probability given the lowest level of executive constraints for
differing values of GDP per capita below the 10\% quantile and down to
the minimum observed value in the dataset while holding constant the
remaining covariates in the same fashion as the interaction plots in
figure \ref{fig:int}.[^theta_bounds] At the far end of the X-axis,
given a GDP per capita of approximately $1,109$ 2011 US dollars, there
is an 80 percent posterior probability that the difference lies within
the interval `r last(df$interval)`. At the opposite end, given a GDP
per capita of approximately $223$ 2011 US dollars, there is an 80
percent posterior probability that the difference is between
`r first(df$interval)`. The lower bounds of the credible intervals remain
fairly consistent across changing values of GDP per capita while the
upper bounds increase due to the distributions of predicted
probabilities being progressively skewed as GDP per capita
decreases. The high degree of uncertainty precludes the ability to
conclusively assert that there is a positive relationship between
civil conflict risk and executive constraints for low income
country-years. This is further bolstered by the fact that there are
relatively few observations falling within the extreme tail ends of
either the executive constraints posterior distribution or the
empirical distribution of GDP per capita --- in the case of the
latter, only `r fmt(mean(X[, "rgdpepc"] < mean(df$y)) * 100)`
percent of the observations lay below the midpoint of the X-axis at
600 2011 US dollars.

[^theta_bounds]: Since the executive constraints parameter, $\theta$,
    is on the unbounded scale, the highest and lowest levels used in
    the calculations for figure \ref{fig:gdp} are drawn from the
    minimum and maximum of the median posterior estimates.

### Assessing Model Fit

(ref:phat) Density plot of predicted probabilities for actual onset
    country-years and no onset country-years.

```{r fig.height = 2, fig.cap = "\\label{fig:phat} (ref:phat)"}
extra <- read_post("posteriors/full_model/extra_posteriors.csv")
p_hat <- extract(extra, "p_hat")

groups <- ifelse(final.df$lepisode_onset == 1,
                 "Actual Onset",
                 "No Onset")
plot_dens(p_hat, groups = groups) +
    xlim(c(0, 0.4))
```

(ref:psum) Summarized posterior predicted probabilities

```{r psum}
p_median <- apply(p_hat, 2, median)

df <- bind_rows(summary(p_median[final.df$lepisode_onset == 1]),
                summary(p_median[final.df$lepisode_onset == 0])) %>%
    mutate_all(~signif(., 3))

rownames(df) <- c("Conflict Onset", "No Onset")

kable(df, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "(ref:psum)")
```

(ref:confusion) Confusion matrix with a 0.168 cutoff for the median
    posterior predicted probabilities.

```{r confusion}
m <- table(p_median > .168, final.df$lepisode_onset)

rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
colnames(m) <- c("Actual: No Onset", "Actual: Onset")

tpr <- fmt(100 * m["Predicted: Onset", "Actual: Onset"] /
              sum(m[, "Actual: Onset"]))
tnr <- fmt(100 * m["Predicted: No Onset", "Actual: No Onset"] /
              sum(m[, "Actual: No Onset"]))

fpr <- fmt(100 * m["Predicted: Onset", "Actual: No Onset"] /
              sum(m[, "Actual: No Onset"]))
fnr <- fmt(100 * m["Predicted: No Onset", "Actual: Onset"] /
              sum(m[, "Actual: Onset"]))

kable(m, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "(ref:confusion)")
```

To assess how well the presented model fits the observed data, I first
consider the predicted probabilities of conflict onset by plotting the
first 50 posterior draws for two groups, country-years that actually
experienced an onset of civil conflict and country-years that did
not. On average, the predicted probabilities for the actual onset
country-years are higher albeit still extremely low. This can be seen
in the summary statistics reported in table \ref{tab:psum} for the
posterior medians of the predicted probabilities. For onset
country-years the interquartile interval is only `r df[1, "1st Qu."]`
to `r df[1, "3rd Qu."]` with a max of `r df[1, "Max."]`. To better
compare the predicted probabilities to the actual data, table
\ref{tab:confusion} presents a confusion matrix with a 0.168 cutoff,
which corresponds to the third quartile in the summary table. `r tpr`
percent of onsets and `r tnr` percent of non-onsets are correctly
predicted. The false positive rate is `r fpr` percent and the false
negative rate is `r fnr` percent.

```{r results = "hide"}
x <- evalmod(scores = p_median, labels = final.df$lepisode_onset)

auc <- attr(x$roc[[1]], "auc") %>% fmt
auc_pr <- attr(x$prcs[[1]], "auc") %>% fmt
```

Figure \ref{fig:roc} plots a receiver operating characteristic (ROC)
curve depicting the tradeoff between the true positive rate
(*sensitivity*) and the false positive rate --- denoted by the formula
$1 - \mbox{specificity}$, where *specificity* refers to the true
negative rate --- for varying probability thresholds according to the
posterior medians of the predicted probabilities. For example, given a
threshold of approximately 0.085, the model will correctly predict
roughly 50 percent of onsets while misclassifying only 8 percent of
non-onsets. The total performance of the model can be summarized by
the area under the curve (AUC), which in this case is `r auc`.[^auc]
However, for the present analysis the results from the ROC curve are
misleading since there is a significant class imbalance in the
dataset. Onsets comprise only
`r fmt(100 * mean(final.df$lepisode_onset == 1))` percent of all
observations meaning that the false positive rate is less sensitive to
changing thresholds. Figure \ref{fig:roc} therefore also plots a
precision-recall (PRC) curve, which depicts the tradeoff between
*precision* --- the number of true positives divided by the sum of the
true positives and false positives --- and *recall* --- the number of
true positives divided by the sum of the true positives and false
negatives --- for varying thresholds. The advantage is that
*precision* is invariant to the number of non-onset years in the
dataset. From this plot it is clear that the precision of the model,
the ability to discriminate between onset country-years and falsely
predicted non-onset country-years, drops drastically as the *recall* of
the model increases. The *AUC-PR* for this plot is only
`r auc_pr`.[^auc_pr]

[^auc]: An AUC of 1 for a ROC curve indicates that the model is able
    to perfectly predict the outcome while an AUC of 0.5 denotes the
    performance of a purely random classifier.

[^auc_pr]: Similar to the ROC curve, an AUC of 1 for a PRC curve
    denotes a perfect model. The AUC for the null model, or random
    classifier, is the proportion of actual positive outcomes to
    negative outcomes in the dataset. For the present model this is
    approximately
    `r fmt(sum(final.df$lepisode_onset == 1) / sum(final.df$lepisode_onset == 0))`.

(ref:roc) ROC & Precision-Recall plots

```{r fig.width = 8, fig.cap = "\\label{fig:roc} (ref:roc)"}
autoplot(x, curvetype = c("ROC", "PRC"))
```

Taken together, the evidence suggests a fairly poor fit to the
underlying data. The confusion matrix demonstrates that even with a
relatively low cutoff, the model is still only able to accurately
predict `r tpr` percent of conflict onsets. According to the ROC
curve, decreasing this threshold improves the true positive rate with
a relatively decent tradeoff in terms of the false positive rate, but
as stated, this is largely due to the imbalance in the outcome
variable. In terms of *precision*, the performance of the model
quickly degrades as the threshold decreases.

## Alternative Specifications

An additional issue which may affect the validity of the model is the
potential sensitivity of the results to changing operational
definitions of civil conflict. Operationalizing civil conflict onset
is a difficult endeavour [@Sambanis2004b]. For example, when exactly
does a conflict start? Is it at the initial formation of a rebel
group? The first battle death? Or, when a certain threshold in
conflict intensity has been achieved? Similar questions can be raised
about when a conflict ends and how to clearly distinguish conflict
termination from brief lulls in fighting. Different operational
definitions can affect estimates and hinder the ability to compare
results across studies. While datasets such as UCDP/PRIO have done
much to standardise the operational definition of civil conflict,
there is still a significant degree of researcher freedom in how onset
is included as the outcome variable in a model.

For the results presented so far the outcome variable has followed the
episodic definition of civil conflict onset from UCDP/PRIO --- 25
battle death threshold with new episodes coded after one year of
inactivity --- which captures both high and low intensity civil
conflicts. However, it is unclear if these findings will remain robust
when only examing full-scale civil wars. There re no *a prior*
theoretical reasons to expect that the relationship between executive
constraints and civil conflict onset should not be negative when only
examining civil wars. In fact, at least one of the mentioned
theoretical mechanisms, namely that executive constraints may affect
the ability of rebel groups to mobilize potential fighters by offering
local security guarantees against government repression, posited that
the theory may *not* hold for low intensity conflicts featuring
predatory rebel groups. Thus, if the relationship were not to be found
when re-running the model with only high intensity conflict onsets
that would be a significant challenge to the external validity of the
initial results.

```{r}
load("./posteriors/cumulative_only/input.RData",
     envir = cumulative <- new.env())

load("./posteriors/major_only/input.RData",
     envir = major <- new.env())
```

In the UCDP/PRIO armed conflict dataset, high intensity is defined at
the threshold of 1,000 battle deaths. High intensity conflicts, or
civil wars, can be coded according to when the threshold for high
intensity is crossed either by a single year or by the cumulative
battle death count.[^battle_deaths] I therefore re-run the same model,
including the same priors and runtime parameters, twice for the two
different operational definitions. In both cases, I retain an episodic
delineation between separate conflicts --- in other words, conflicts
over the same incompatibility can have multiple onsets when there are
lulls in fighting.[^episode_gaps] For the former, the onset year is
set as the first year which crosses the high intensity threshold for
each period of sustained fighting of at least 25 battle deaths. For
the latter, onset is the first coded conflict year for the cumulative
period passing the high intensity threshold. In total there are
`r sum(major$final.df$lmajor_onset)` onsets for the single year 1,000
battle deaths definition, and `r sum(cumulative$final.df$lcum_onset)`
onsets for the cumulative definition.

[^battle_deaths]: A more robust modeling strategy would be to directly
    predict the number of battle deaths and then apply *a posteriori*
    decision criteria to the predicted results to classify
    country-years as low-intensity or high intensity conflict. This
    would avoid the arbitrary nature of battle deaths thresholds for
    determining the intensity types; however, battle death data tends
    to be limited in time span and there would be the additional issue
    of data quality.

[^episode_gaps]: Another robustness check would be to re-run the model
    for different episode definitions based on varying intervals of
    fighting inactivity. For example, coding termination after one,
    two, or three years of inactivity. This remains outside the scope
    of the present paper, however, due to lack of computational
    resources for such a high number of model runs.

(ref:overlay) Comparison of regression slopes between the original
    model run, "Low & High Intensity", and the model runs for the two
    different operational definitions of high intensity
    conflict. Credible intervals are 80 percent posterior quantile
    intervals.

```{r fig.width = 8, fig.height = 4, fig.cap = "\\label{fig:overlay} (ref:overlay)"}
models <- c("full_model", "major_only", "cumulative_only")
ll <- lapply(models, function(model) {
    post <- file.path("posteriors", model, "reg_posteriors.csv") %>%
        read_post
})

args <- list(pars = "beta",
             legend_labels = c("Low & High Intensity",
                               "High Intensity",
                               "Cumulative High Intensity"))
do.call(overlay_pars, c(ll, args)) +
    scale_x_discrete(labels = colnames(beta))
```

```{r}
ix_probs <- lapply(seq_along(models), function(i) {
    beta <- extract(ll[[i]], "beta")

    if (models[i] == "major_only") {
        colnames(beta) <- c("exec_constraints", colnames(major$X),
                            sprintf("exec * %s",
                                    colnames(major$X)[major$interaction_idx]))
    } else {
        colnames(beta) <- c("exec_constraints", colnames(cumulative$X),
                            sprintf("exec * %s",
                                    colnames(cumulative$X)[cumulative$interaction_idx]))
    }

    mean(beta[, "exec * rgdpepc"] < 0)
})

ix_probs <- unlist(ix_probs)
names(ix_probs) <- models
```

Figure \ref{fig:overlay} plots the original regression slopes compared
to the results from running the model with the two different
operational definitions of high intensity conflict.[^theta] The
majority of depicted parameter estimates are fairly stable across
runs. Notably, the relationship between executive constraints and the
risk of civil conflict onsets remains negative with a high posterior
probability of being less than zero. The estimates for the interaction
effect between executive constraints and GDP per capita, however,
become far too uncertain to be able to conclusively assert an
association. The probability that the interaction term is less than
zero for the high intensity model run, *i.e.* where at least a single
year crosses the high intensity threshold, is only `r
fmt(ix_probs["major_only"])` while the probability for the cumulative
intensity model run is only `r fmt(ix_probs["cumulative_only"])`.

[^theta]: Since the measurement model for executive constraints is
    embedded within the larger regression predicting civil conflict
    onset, the latent factor $\theta$ is re-estimated for each model
    run. The stochastic nature of posterior sampling ensures that the
    estimates are never exactly the same; however, the differences are
    inconsequential. In fact the correlation between the main model
    and both the cumulatively high intensity model and the major
    conflicts only model is effectively one.

(ref:confusion-major) Confusion matrix with a 0.168 cutoff for the
    median posterior predicted probabilities from the single year high
    intensity model run.

```{r confusion-major}
extra <- read_post("./posteriors/major_only/extra_posteriors.csv")

p_hat <- extract(extra, "p_hat")
p_median <- apply(p_hat, 2, median)

m <- table(p_median > .168, major$final.df$lmajor_onset)
rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
colnames(m) <- c("Actual: No Onset", "Actual: Onset")

# Calculate the AUCs as well
x <- evalmod(scores = p_median, labels = major$final.df$lmajor_onset)
major_auc <- attr(x$roc[[1]], "auc") %>% fmt
major_auc_pr <- attr(x$prcs[[1]], "auc") %>% fmt

kable(m, "latex", align = "c", booktabs = T,
      linesep = "", escape = F, caption = "(ref:confusion-major)")
```

Table \ref{tab:confusion-major} presents the confusion matrix for the
single year high intensity model run using the median posterior
predicted probabilities. A cutoff of 0.168 was reused for comparative
purposes to the original results. In this case, the model is unable to
predict even a single major conflict onset with the highest median
posterior predicted probability for an actual onset being only
`r fmt(max(p_median[major$final.df$lmajor_onset == 1]))`. Furthermore,
although the ROC AUC is `r major_auc`, this is again due to the major
class imbalance between positive versus negative outcomes in the
dataset. A precision-recall AUC of `r major_auc_pr` demonstrates the
poor performance of the model.

(ref:confusion-cumulative) Confusion matrix with a 0.168 cutoff for
    the median posterior predicted probabilities from the cumulative
    high intensity model run.

```{r confusion-cumulative}
extra <- read_post("./posteriors/cumulative_only/extra_posteriors.csv")

p_hat <- extract(extra, "p_hat")
p_median <- apply(p_hat, 2, median)

m <- table(p_median > .168, cumulative$final.df$lcum_onset)
rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
colnames(m) <- c("Actual: No Onset", "Actual: Onset")

x <- evalmod(scores = p_median, labels = cumulative$final.df$lcum_onset)
cum_auc <- attr(x$roc[[1]], "auc") %>% fmt
cum_auc_pr <- attr(x$prcs[[1]], "auc") %>% fmt

kable(m, "latex", align = "c", booktabs = T,
      linesep = "", escape = F,
      caption = "(ref:confusion-cumulative)")
```

Similarly, table \ref{tab:confusion-cumulative} shows the confusion
matrix for the cumulative high intensity model run with the same
median posterior predicted probability cutoff. Although the model in
this case is able to correctly classify 17 percent of onsets with an
ROC AUC of `r cum_auc`, the overall performance is fairly limited with
a precision-recall AUC of only `r cum_auc_pr` out of 1.

Taken together the model runs examining high intensity conflicts
suggest that the negative relationship between executive constraints
and civil conflict onset is fairly robust. However, these model runs
similarly suffer from the same issues of relatively poor model fit.

# Conclusion {-}

How do the presented results from the various model runs relate back
to the theories potentially linking executive constraints to civil
conflict? According to the primary model, as well as the robustness
checks, there is support for the general proposition that the risk of
civil conflict onset decreases the more constrained the executive is
by other government branches when controlling for the selected
covariates. As stated previously, however, an in-depth investigation
of which theoretical mechanism(s) might be reponsible for this
relationship lies outside of the scope for this paper in large part
because the mechanisms that were discussed in the theory section are
not mutually exclusive.  Furthermore, quantitative country-year
regressions such as the ones performaned in this thesis can only
detect general patterns and provide evidence towards the associations
between particular indicators. What regression models cannot do is
examine the microfoundations underpinning a particular theory.

That being said, on the basis of the mobilization theory I argued that
there may be a potential confounder in the relative threshold for
mobilization. I operationalized this in the model as GDP per capita
and included an interaction term between GDP per capita and the latent
construct for executive constraints. Accordingly, the results show
that the negative relationship between executive constraints and civil
conflict onset does not hold for low-income country-years. In these
cases, the effect is practically non-existent. According to theory,
this would mean that where labor costs are low and the provisioning of
economic selective benefits presents less of an obstacle for rebel
leaders to mobilize fighters, the effect of executive constraints on
the risk of civil conflict is largely indeterminate. This may be
because mobilization in these situations is independent of the
government's decision to engage in repression, assuming that the state
does not have the capacity to completely eradicate the opposition
immediately. More specifically, low-income countries may be vulnerable
to the types of predatory insurgent groups which require significantly
less fighters than rebel groups seeking regional independence or to
topple the central regime. This interpretation could also hold even if
GDP per capita was instead understood as proxying the state's
administrative and military capacity by roughly capturing the revenue
potential available to the state [@Fearon2003]. Where the state is
particularly weak it simply might not matter how constrained the
executive is.

These results, however, should be interpreted with caution as there is
a significant caveat to the presented model. Namely, the model and the
estimated predicted probabilities were a poor fit to the underlying
data when examining both low intensity and high intensity civil
conflicts. The tradeoff between precision and recall was far from
satisfactory; the model could easily predict non-onset country-years
in large part because of the class imbalance in the dataset between
onsets and non-onsets, but had difficultly predicting actual onsets
with accuracy. This is unsurprising given that civil conflict onsets
are relatively rare events and studies attempting to examine the
impact of structural factors, such as political institutions, on onset
risk are dealing with small effect sizes, noisy data, and relatively
few variables of actual exogenous change. This is an issue that has
plagued much of the conflict literature [@Ward2010] so the model
presented here is far from being an outlier. But, it does mean that any
results should be considered tentative at best, at least until
confirmed by additional studies.

Finally, in terms of future research, if we want to better understand
the impact of political institutions on the risk of civil conflict we
need to directly model the behaviour of rebel groups and state
actors. For executive constraints, and specifically the theoretical
mechanism focused on rebel mobilization, one possible avenue that was
unexplored in this paper would be to model the effects of
repression. A future study could examine the temporal sequence in the
development of executive constraints on the likelihood of government
repression and its effect in turn on the risk of civil conflict
onset. Such an analysis would move beyond the standard country-year
logistic regression, which is a poor model for the dynamics of civil
conflict, and attempt to model country-date data, possibly even at the
subnational level depending on data availability.

# Appendices {-}

## A - Stan Code {#appendix-a}

```{stan, code = readLines("stan/sem.stan"), eval = F, output.var = "x", echo = T}
```

\newpage

## B - Prior Simulation {#appendix-b}

In order

```{r eval = F}
prior_plot <- function(range, fn, title = NULL, ...) {
    data.frame(x = range) %>%
        ggplot(aes(x)) +
        stat_function(fun = fn, ...) +
        ggtitle(label = title) +
        theme(axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank())
}

p1 <- prior_plot(c(-20, 20), dnorm, args = list(mean = 0, sd = 5),
                 title = "Normal(0, 5)")
p2 <- prior_plot(c(-10, 10), dnorm, args = list(mean = 0, sd = 2.5),
                 title = "Normal(0, 2.5)")
p3 <- prior_plot(c(0, 5), dlnorm, args = list(meanlog = 0, sdlog = 0.5),
                 title = "Log-Normal(0, 0.5)")
p4 <- prior_plot(c(0, 3), dweibull, args = list(shape = 5, scale = 1),
                 title = "Weibull(5, 1)")
p5 <- prior_plot(c(0, 10), dhcauchy, args = list(sigma = 1),
                 title = "HalfCauchy(0, 1)")
p6 <- prior_plot(c(0, 1), dbeta, args = list(shape1 = 5, shape2 = 5),
                 title = "Beta(5, 5)")

grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 2, nrow = 3)

```

```{r eval = F, warning = F}
true_values <- readRDS("posteriors/sim/simulated_parameters.rds")

fa <- read_post("posteriors/sim/fa_posteriors.csv")
reg <- read_post("posteriors/sim/reg_posteriors.csv")

fit <- bind_cols(fa, reg)

post <- post_summarise(fit, c("lambda", "psi", "gamma", "delta",
                              "kappa", "rho", "eta", "alpha", "beta",
                              "sigma")) %>%
    rename(codelow = `2.5%`, median = `50%`, codehigh = `97.5%`) %>%
    mutate(type = "Posterior Estimate")

pars.df <- filter(true_values, !grepl("theta", parameter)) %>%
    bind_rows(post) %>%
    mutate(type = as.factor(type))

ggplot(pars.df, aes(parameter, point, color = type)) +
    geom_point() +
    geom_errorbar(aes(parameter, ymin = codelow, ymax = codehigh),
                  width = 0) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r eval = F, warning = F}
latent.df <- post_summarise(fit, pars = "theta", probs = c(0.025, 0.975)) %>%
    rename(codelow = `2.5%`, codehigh = `97.5`)

x <- mean(theta <= latent.df$codehigh & theta >= latent.df$codelow)

```

```

\newpage

# References {-}

\singlespacing